---
title: "data_analysis"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(factoextra)
```

### Import data

```{r}
pmd <- read.csv("../../data/pmd-en.csv")
head(pmd, 3)
```

### DATA Summary

```{r}
# no. of rows and columns
dim(pmd)
```

489676 rows and 41 columns

```{r}
# dataset summary
str(pmd)
```


```{r}
# percentage of missing values in each column in dataset
p <- function(x) {sum(is.na(x))/length(x)*100}
sort(apply(pmd, 2, p), decreasing = TRUE) # marmin = 2 means function will be applied in each column 
```

In head of dataset we saw there were missing values in `prox_idx_lib` but the above output suggest there is no missing values. Because Statistics Canada use some specific notation for missing values. The following standard symbols are used in Statistics Canada publications:  
$.. ->$ not available for a specific reference period  
$F ->$ to unreliable to be published

```{r}
# percentage of missing values in each column
p <- function(x) {sum(x == ".." | x == "F")/length(x)*100}
sort(apply(pmd, 2, p), decreasing = TRUE)
```

```{r}
# Remove rows with ".." or "F" values in proximity columns
pmd_filtered <- pmd %>%
  filter_at(vars(starts_with("prox_idx")), all_vars(. != ".." & . != "F"))
```

```{r}
# no. of rows and columns in filtered dataset
dim(pmd_filtered)
```


```{r}
p <- function(x) {sum(x == ".." | x == "F")/length(x)*100}
sort(apply(pmd_filtered, 2, p), decreasing = TRUE)
```

```{r}
# Convert selected columns to numeric
pmd_filtered <- pmd_filtered %>%
  mutate(prox_idx_emp = as.numeric(prox_idx_emp),
         prox_idx_pharma = as.numeric(prox_idx_pharma),
         prox_idx_childcare = as.numeric(prox_idx_childcare),
         prox_idx_health = as.numeric(prox_idx_health),
         prox_idx_grocery = as.numeric(prox_idx_grocery),
         prox_idx_educpri = as.numeric(prox_idx_educpri),
         prox_idx_educsec = as.numeric(prox_idx_educsec),
         prox_idx_lib = as.numeric(prox_idx_lib),
         prox_idx_parks = as.numeric(prox_idx_parks),
         prox_idx_transit = as.numeric(prox_idx_transit))
```

```{r}
# Remove specified columns from dataset
df_filtered <- select(pmd_filtered, -DBUID, -DBPOP, -DAUID, -DAPOP, -CSDUID, -CSDNAME, -CMAPUID,
                      -CSDTYPE, -CSDPOP, -CMANAME, -CMATYPE, -CMAPOP, -PRUID, -PRNAME, -PRPOP, -CMAUID, -in_db_emp, -in_db_pharma, -in_db_childcare, -in_db_health, -in_db_grocery, -in_db_educpri, -in_db_educsec, -in_db_lib, -in_db_parks, -in_db_transit, -transit_na, -amenity_dense, -suppressed)
```

```{r}
any(is.na(df_filtered))
```
```{r}
head(df_filtered)
```
```{r}
df_filtered <- select(df_filtered, -lon, -lat)
```

```{r}
set.seed(123)
km.res <- kmeans(df_filtered, 4, nstart = 25)
```

```{r}
# Print the results
print(km.res)
```

```{r}
dd <- cbind(df_filtered, cluster = km.res$cluster)
```

```{r}
dd[, c("cluster")]
```

```{r}
# Visualize the clusters
fviz_cluster(km.res, data = df_filtered, geom = "point", ellipse.type = "convex")
```

```{r}
str(df_filtered)
```


```{r}
fviz_cluster(res.km, data = df[, -5],
             palette = c("#2E9FDF", "#00AFBB", "#E7B800"), 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
             )
```











