---
title: "data_analysis"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T)
library(dplyr)
library(factoextra)
library(ggplot2)
library(tidyverse)
```


### Import data

```{r}
pmd <- read.csv("../../data/pmd-en.csv")
head(pmd, 3)
```

### DATA Summary

```{r}
# no. of rows and columns
dim(pmd)
```

489676 rows and 41 columns

```{r}
# dataset summary
str(pmd)
```

```{r}
# definitions of variables
defpmd <- read.csv("../../data/pmd-en-record-layout-notes.csv")
```

Definitions: https://www150.statcan.gc.ca/n1/pub/92f0138m/92f0138m2019001-eng.htm



In head of dataset we saw there were missing values in `prox_idx_lib` but the above output suggest there is no missing values. Because Statistics Canada use some specific notation for missing values. The following standard symbols are used in Statistics Canada publications:  
$.. ->$ not available for a specific reference period  
$F ->$ to unreliable to be published (data sources disagree)

```{r}
# changing '..' and 'F' to Nas
pmd_na <- readr::read_csv("../../data/pmd-en.csv", na = c('..', 'F'))
summary(pmd_na)
```
```{r}
# getting proportions of nas
na_prop <- function(x) {sum(is.na(x))/length(x)*100}
sort(apply(pmd_na, 2, na_prop), decreasing = T)
```


Why would we be missing information about population? 
```{r}
table(filter(pmd_na, is.na(DBPOP))$CSDTYPE)
```
IRI – Indian reserve
S-É – Indian settlement 


# Missing data analysis

Trying to make the dataset smaller by conducting analysis by province.

I want to do a spatial analysis to see if the NA values are correlated with things like population.Assume population is a 'continuous' covariate, maybe pull in remoteness index as well as covariate. See if 'point process' of missing values is correlated. Could see if could add information about gdp, need to merge with other datasets I guess... Right now: want to plot the missing value of grocery locations on a map of canada, for BC. 

```{r}
# only grocery NAs and BC
groc_na <- filter(pmd_na, is.na(pmd_na$prox_idx_grocery) & pmd_na$PRUID == 59) # for BC
#groc_na <- filter(pmd_na, is.na(pmd_na$prox_idx_grocery))
```

```{r message=F}
library(sf)
library(sp) 
library(spatstat)
library(maptools)
```

```{r}
# Create a spatial points data frame from the longitude and latitude columns
coordinates <- groc_na[,c("lon", "lat")]
dat.sp <- SpatialPointsDataFrame(c(groc_na[,c("lon", "lat")]), data = groc_na)

# Set the current CRS: assume it is the same as from shapefile, std canadian ?
proj4string(dat.sp)<- CRS("+proj=longlat +datum=WGS84")
#proj4string(dat.sp)<- CRS("+proj=lcc +lat_0=63.390675 +lon_0=-91.8666666666667 +lat_1=49 +lat_2=77 +x_0=6200000 +y_0=3000000 +datum=NAD83 +units=m +no_defs")

##  Define the new CRS you want to transform to
# new_crs <- CRS("+proj=aea +lat_0=45 +lon_0=-126 +lat_1=50 +lat_2=58.5 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs")
new_crs <- CRS("+proj=lcc +lat_0=63.390675 +lon_0=-91.8666666666667 +lat_1=49 +lat_2=77 +x_0=6200000 +y_0=3000000 +datum=NAD83 +units=m +no_defs")

# Transform the data to the new CRS
data.sp_trans <- spTransform(dat.sp, new_crs)
```

```{r eval = F}
# importing for the province, for nicer plotting?
provshp <- readShapeSpatial("../../data/prov_g/lpr_000b21a_e.shp")
provshp <- as(provshp, "SpatialPolygons")
provshp <- as.owin(provshp)
saveRDS(provshp, file = "prov_owin.rds")
plot(provshp)
```

```{r}
provshp <- readRDS("prov_owin.rds")
```


```{r eval=F}
# importing the shapefile for each DB
dbshp <- readShapeSpatial("../../data/DB/ldb_000a21a_e.shp")
dbshp <- as(dbshp, "SpatialPolygons")
dbshp <- as.owin(dbshp)
saveRDS(dbshp, file = "dbshp_owin")
plot(dbshp)
```


```{r}
groc_ppp <- ppp(x = data.sp_trans@coords[,1], # X coordinates
                    y = data.sp_trans@coords[,2], # Y coordinates
                    window = provshp,# Observation window, but from prov_g file
                )

plot(groc_ppp)
```
```{r}
saveRDS(groc_ppp, file = "groc_ppp.rds")
```
```{r}
plot(groc_ppp)
```

```{r}
# this is when I set dat.sp to have the std canadian CRS, apparently none of them plot... 
groc_ppp2 <- ppp(x = dat.sp@coords[,1], # X coordinates
                    y = dat.sp@coords[,2], # Y coordinates
                    window = provshp,# Observation window, but from prov_g file
                )
```


```{r}
# finding out what the proj string is
library(rgdal)
# read the .shp file - layer is the same name but without the .shp
mymap <- readOGR("../../data/prov_g/lpr_000b21a_e.shp", layer="lpr_000b21a_e") 
# proj4 info is located in the layer's proj4string slot
mymap@proj4string

"+proj=lcc +lat_0=63.390675 +lon_0=-91.8666666666667
+lat_1=49 +lat_2=77 +x_0=6200000 +y_0=3000000
+datum=NAD83 +units=m +no_defs "
"+proj=lcc +lat_0=63.390675 +lon_0=-91.8666666666667
+lat_1=49 +lat_2=77 +x_0=6200000 +y_0=3000000
+datum=NAD83 +units=m +no_defs"
```


options(maxppsize = 500000)
```{r}
saveRDS(groc_ppp, file = "groc_ppp.rds")
```

```{r}
df_ppp <- readRDS("../../../df_ppp.rds")
```


```{r}
plot(df_ppp)
```

```{r}
Q10 <- quadratcount(df_ppp,
                  nx = 10,
                  ny = 10)
```
# doing onto BC window from 589
```{r}
load("~/Desktop/589/589Project/BC_Covariates.Rda")

bc_crs <- CRS("+proj=aea +lat_0=45 +lon_0=-126 +lat_1=50 +lat_2=58.5 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs")

# Transform the data to the new CRS
data.bc <- spTransform(dat.sp, bc_crs)



groc_ppp_bc <- ppp(x = data.bc@coords[,1], # X coordinates
                    y = data.bc@coords[,2], # Y coordinates
                    window = as.owin( DATA[["Window"]]),# Observation window from BC_covariates RDS
                )

plot(groc_ppp_bc)
```


# Proportion of missing values by amenity and province
See what proportion of missing values for each province, if it is proportion for total number of DB for each province

I got this from manually comparing the prov codes and the province name:
AB: 48
BC: 59
MN: 46
NB: 13
NL: 10
NWT: 61
NC: 12
NV: 622
ON: 35
PEI: 11
QB: 24
SK: 47
YK: 60

```{r}
# Getting infos from data 
amen_cols <- names(select(pmd_na, starts_with("prox"))) # columns names of proximity measures
num_amen_cols <- grep("prox", colnames(pmd_na)) # column numbers of proximity measures
prov_code <- c(48,59,46,13,10,61,12,62,35,11,24,47,60) # province codes
prov_name <- c("AB", "BC", "MN", "NB","NL", "NT", "NC", "NU","ON", "PEI", "QB","SK","YK" ) # province abbreviations


#initialize a matrix and empty list
mat_missing = matrix(data=NA, ncol=length(num_amen_cols), nrow=length(prov_code))
amt_rows <- c()

# Making a matrix of amenities x provinces, percentage missing values
for (i in 1:length(prov_code)){
  # get the number of rows for each province
  amt_rows[i] <- nrow(filter(pmd_na, pmd_na$PRUID == prov_code[i]))
  for (j in 1:length(num_amen_cols)){
    # for every province, iterate through every amenity and get percentage missing
    amt_miss <- nrow(filter(pmd_na, is.na(pmd_na[,num_amen_cols[j]]) & pmd_na$PRUID == prov_code[i]))
    mat_missing[i,j] <- amt_miss/amt_rows[i]
  }
}

# build dataframe, rename columns and rows
df_perc_missing <- mat_missing %>% data.frame() %>% setNames(amen_cols)
rownames(df_perc_missing) <- prov_name

df_perc_missing
```

# Idea: build logistic model to see if can predict if missing or not based on population or other variables
```{r}
# convert data so NA = 1, other = 0
df_na <- pmd_na

# converting all values to 0 and all na to 1 for amenities cols
df_na[amen_cols][!is.na(df_na[amen_cols])] <- 0
df_na[amen_cols][is.na(df_na[amen_cols])] <- 1

log_mod <- glm(prox_idx_emp~DBPOP, data = df_na)
summary(log_mod)
plot(log_mod)
```
```{r}
# plot logistics curve for employment and population
ggplot(df_na, aes(x=DBPOP, y=prox_idx_emp)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
```

```{r}
# plot histogram
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_emp)), bins = 50,
                 colour="grey50", alpha=0.5, position="identity")
```


# Idea: look at distribution of proximity measures


```{r}
# manipulate dataframe to only keep amenities cols
df_man <- subset(pmd_na, select = amen_cols)

# pivot to long to make work with violin plot
df_long <- pivot_longer(df_man, all_of(amen_cols))
# whole range
ggplot(df_long, aes(x=name, y = value)) + geom_violin()

# setting y lim to see better
ggplot(df_long, aes(x=name, y = value))+ geom_violin()  + ylim(0,0.25)
```
When plotting everything on the same plot, it is hard to see. For all amenities, the highest densities of proximity values lie below 0.12. We see that the amenities with the highest distribution density closer to 0 are health, then employment, then transit. 

## Histograms
```{r}
hist(pmd_na$prox_idx_lib, breaks = 50)
```

```{r fig.height=10, fig.width=12}
# histograms
ggplot(df_long, aes(value)) + 
    geom_histogram(bins = 50) + 
    facet_wrap(~name, scales = 'free_y', ncol = 2)
```
```{r fig.height=10, fig.width=12}
# kernel densities
ggplot(df_long, aes(value)) + 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 2)
```


## plot 5

```{r}
# plotting first 5 then next 5
# pivot to long to make work with violin plot
df_long_5 <- pivot_longer(df_man, all_of(amen_cols[1:5]))
# first half
ggplot(df_long_5, aes(x=name, y = value)) + geom_violin()
```


```{r}
# pivot to long to make work with violin plot
df_long_10 <- pivot_longer(df_man, all_of(amen_cols[6:10]))
# 2nd half
ggplot(df_long_10, aes(x=name, y = value)) + geom_violin()
```
## plot 2
```{r}
# pivot to long to make work with violin plot
df_long_2 <- pivot_longer(df_man, all_of(amen_cols[1:2]))
df_long_4 <- pivot_longer(df_man, all_of(amen_cols[3:4]))
df_long_6 <- pivot_longer(df_man, all_of(amen_cols[5:6]))
df_long_8 <- pivot_longer(df_man, all_of(amen_cols[7:8]))
df_long_10_2 <- pivot_longer(df_man, all_of(amen_cols[9:10]))
# 2nd half
ggplot(df_long_2, aes(x=name, y = value)) + geom_violin()
```
It looks like for both employment and pharmacy proximity measure, most are very small: both have the shape of a funnel. There are a lot more values close to 0 for employment proximity, and it looks almost like an exponential decrease as the value increase, whereas the rate of decrease for pharma is smaller. 

```{r}
ggplot(df_long_4, aes(x=name, y = value)) + geom_violin()
```


```{r}
ggplot(df_long_6, aes(x=name, y = value)) + geom_violin()
```
```{r}
ggplot(df_long_8, aes(x=name, y = value)) + geom_violin()
```

```{r}
ggplot(df_long_10_2, aes(x=name, y = value)) + geom_violin()
```
