---
title: "EDA draft"
author: "Ricky Heinrich & Team"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = T, message = F, warning = F, fig.height=4, fig.width=6)
library(dplyr)
library(factoextra)
library(ggplot2)
library(tidyverse)
library(gridExtra)
```

## Outline
- Overall summary: how many variables, how many missing values

- Exploring missing values: are they related to population?

- Exploring data distributions

- Population zero


# Overall Summary
```{r message = F, warning = F}
#importing data and changing '..' and 'F' to Nas
pmd_na <- readr::read_csv("../../data/pmd-en.csv", na = c('..', 'F'))
# dim(pmd_na)
```
There are 489 676 rows in this data and 41 columns, meaning that 489 676 dissemination blocks are included. The 41 columns include information about the dissemination blocks themselves such as ID, population, and coordinates, as well as information about other census boundaries like dissemination areas, census areas, and provinces. Each of the 10 amenities have two columns associated with it: one a binary indicator to track whether the amenity is present in the DB itself, and the other the calculated proximity measure. Finally there are three indicators: transit_na, amenity_dense, and suppressed.

In the summary of the dataset, we see that there are many missing values. We see that the library proximity indicator contains the most missing values, at around 77%, followed by the proximity measures for grocery and secondary education. Only two out of the ten amenities have proximity measures missing proportion under 50%: health and employment. 

```{r}
# getting proportions of nas
na_prop <- function(x) {sum(is.na(x))/length(x)*100}
sort(apply(pmd_na, 2, na_prop), decreasing = T)
```

In the following chart, can see the proportion of missing values for each amenity by province. We see that overall, employment has the lowest rates of missing values, but has also more range depending on the province. Ontario and Quebec seems to have the least missing values for most amenities relative to the other regions, whereas Nunavut usually has the most. It seems like the amount of proximity measure missing for libraries are the most consistent across regions. 

```{r}
# Getting infos from data 
amen_cols <- names(select(pmd_na, starts_with("prox"))) # columns names of proximity measures
num_amen_cols <- grep("prox", colnames(pmd_na)) # column numbers of proximity measures
prov_code <- c(48,59,46,13,10,61,12,62,35,11,24,47,60) # province codes
prov_name <- c("AB", "BC", "MN", "NB","NL", "NT", "NC", "NU","ON", "PEI", "QB","SK","YK" ) # province abbreviations


#initialize a matrix and empty list
mat_missing = matrix(data=NA, ncol=length(num_amen_cols), nrow=length(prov_code))
amt_rows <- c()

# Making a matrix of amenities x provinces, percentage missing values
for (i in 1:length(prov_code)){
  # get the number of rows for each province
  amt_rows[i] <- nrow(filter(pmd_na, pmd_na$PRUID == prov_code[i]))
  for (j in 1:length(num_amen_cols)){
    # for every province, iterate through every amenity and get percentage missing
    amt_miss <- nrow(filter(pmd_na, is.na(pmd_na[,num_amen_cols[j]]) & pmd_na$PRUID == prov_code[i]))
    mat_missing[i,j] <- amt_miss/amt_rows[i]
  }
}

# build dataframe, rename columns and rows
df_perc_missing <- mat_missing %>% data.frame() %>% setNames(amen_cols)
rownames(df_perc_missing) <- prov_name

df_perc_missing <- cbind(prov_name, df_perc_missing)

#df_perc_missing
```
```{r fig.height=10, fig.width=12}
# flattening so can facet
df_perc_missing %>% pivot_longer(all_of(amen_cols)) %>%
  ggplot(aes(y = value, x = prov_name, fill = prov_name)) + 
    geom_col() + 
    facet_wrap(~name, ncol = 2) + labs(title = "Missing values for each province by amenity")
```

For each amenity, we can plot the occurrence of missing values in a DB vs its population, and plot a basic logistics curve. The employment curve is included below, and the remainder are found in the appendix. We see that for some amenities, like employment and health, the missing values are concentrated among DBs with small populations. These are the same amenities with less than 50% of values missing. Overall it seems like the population of the DB is not the only factor, if at all, affecting whether a proximity measure is missing for that DB.

```{r}
# convert data so NA = 1, other = 0
df_na <- pmd_na

# converting all values to 0 and all na to 1 for amenities cols
df_na[amen_cols][!is.na(df_na[amen_cols])] <- 0
df_na[amen_cols][is.na(df_na[amen_cols])] <- 1
```

```{r fig.height=3, fig.width=6}
# plot logistics curve for amenity and population
# couldn't figure out easy way to get it; for loops weren't working, and facet involves flattening but I wanted to keep most of the info incase I needed it
#par(mfrow = c(2,5))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_emp)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial)) + labs(title = "Logistic regression curve (1-NA) vs population")

```


We can also plot the histograms of missing values vs populations for each amenity, where '1' (blue) is a missing value and '0' (pink) is a value not missing. Again, the remainder are found in the appendix.

```{r fig.height=4, fig.width=6}
# I should really just facet it
par(mfrow = c(5,2))
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_emp)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity") + labs(title = "Histogram NA values vs rest by population for employment")
```

Overall, there are more DBs with lower populations than larger populations. We see that for some amenities, at smaller populations, there are a lot more missing values. Again, employment and health are the only two where there are always more actual values than missing values at every population bin. 

```{r eval = F}
# ### Model with other variables to see relationship: need add 'master' df
# basic model with only population
log_mod <- glm(prox_idx_emp~DBPOP, data = df_na)
summary(log_mod)
# diagnostics
plot(log_mod)
```


# Data Distributions
We can take a preliminary look at the distribution of proximity measures for each amenity, to see if there are 'obvious' clusters.

```{r}
# manipulate dataframe to only keep amenities cols
df_man <- subset(pmd_na, select = amen_cols)
# rename cols for easier plot reading
names(df_man) <- sub('prox_idx_', '', names(df_man))
amen_cols_short <- names(df_man)
# pivot to long to make work with plots
df_long <- pivot_longer(df_man, all_of(amen_cols_short))
```

In this violin plot, we see that the highest densities of proximity values lie below 0.12 for all amenities. We see that the amenities with the highest distribution density closer to 0 are health, then employment, then transit. Library has the lowest density right around zero, and 'starts' a bit later. Health and employment have the least amount of missing values, and library has the most; some conclusion could be made out of that. 

```{r fig.height= 4, fig.width=6}
ggplot(df_long, aes(x=name, y = value)) + geom_violin() + labs(title = "Distribution of proximity measures by amenity")
```





Next we see the kernel densities of proximity measures for each amenity. We see that most curves appear smooth, but some like for primary education, secondary education, and library, have 'bumps', which could indicate clusters. Overall, the naked eye is not able to perceive obvious segmentation cutoffs.

```{r fig.height=10, fig.width=6}
# kernel densities
ggplot(df_long, aes(value)) + 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 2) + labs(title = "Distribution of proximity measures by amenity")
```


# Population zero

```{r include=F}
# how many DB have population of 0
nrow(filter(pmd_na, pmd_na$DBPOP == 0))/nrow(pmd_na)
```

About 24% of the DBs in Canada (in 2016) have a population of 0. It could be reasonable to expect that if the population of a DB is 0, then the proximity measure are also near 0: it is intuitive that for the most part, amenities are further away from areas with no populations. It is thus reasonable to explore the cases where the population is zero, to see its prevalence, and deduce how it may affect the values of proximity measures. In the appendix there's a barplot showing how many DBs there are per province: Ontario and Quebec have the most, whereas the Territories have the least. 

Here, we that the province with the most DBs with a population of zero is Saskatchewan, followed by Ontario, Quebec, and Alberta. 

```{r}
PRUID <- unique(pmd_na$PRUID)
pr_abbr <- c('NL','PE','NC','NB','QB','ON','MN','SK','AB','BC','YK', 'NW','NV')

pr_df <- cbind(PRUID, pr_abbr)
pmd_na <- merge(pmd_na, pr_df, by = 'PRUID', all.x = T)

pmd_na %>% filter(pmd_na$DBPOP == 0) %>% 
  ggplot(aes(pr_abbr)) + geom_bar() + labs(title = "How many DBs with population zero by province") 

#cowplot::plot_grid(p1,p2)
```

Taking the proportions however, we see that over 50% of Yukon and NWT's DBs have a population of 0, and Saskatchewan has over 40%. Ontario has the lowest at around 15%, followed by Quebec at around 18%.

```{r}
# should get proportions
data.frame(table(filter(pmd_na, pmd_na$DBPOP == 0)$pr_abbr)/table(pmd_na$pr_abbr)) %>% ggplot(aes(factor(Var1), Freq)) + geom_col() + labs(title = "Proportion of population zero by province")
```

In the following table, we see that the CSDTYPE with the most populations = zero are rural municipalities followed relatively closely by cities. It seems the majority of the top counts are urban areas (cities, villes, municipalities, etc), which is somewhat unexpected.

```{r}
sort(table(filter(pmd_na, pmd_na$DBPOP == 0)$CSDTYPE), decreasing = T)
```

Some of the population is NA. We see that Quebec has the most DBs with a population NA, followed by Ontario, Alberta, and BC. The CSDTYPE of the DB's whose population information is NA are IRI – Indian reserve and S-É – Indian settlement. 

```{r}
pmd_na %>% filter(is.na(pmd_na$DBPOP)) %>% ggplot(aes(factor(pr_abbr))) + geom_bar() + labs(title = "How many DBs with population NA by province")
```
```{r eval=F}
table(filter(pmd_na, is.na(DBPOP))$CSDTYPE)
```

## Effect of removing population = 0

```{r}
# Data wrangling
# keeping only vals not 0
pop_0 <- filter(pmd_na, pmd_na$DBPOP !=0)
names(pop_0) <- sub('prox_idx_', '', names(pop_0))
# manipulate dataframe to only keep amenities cols
df_man_pop <- subset(pop_0, select = amen_cols_short)

# pivot to long to make work with plots
df_long_pop <- pivot_longer(df_man_pop, all_of(amen_cols_short))

# if only keep population ==0
pop_only0 <- filter(pmd_na, pmd_na$DBPOP ==0)
names(pop_only0) <- sub('prox_idx_', '', names(pop_only0))
# manipulate dataframe to only keep amenities cols
df_man_pop_only <- subset(pop_only0, select = amen_cols_short)

# pivot to long to make work with plots
df_long_pop_only <- pivot_longer(df_man_pop_only, all_of(amen_cols_short))
```


```{r fig.height=10, fig.width=6}
# data wrangling
# add row indicating what data type
which_df <- rep("all", nrow(df_long))
df_long <- cbind(df_long, which_df)
which_df <- rep("red", nrow(df_long_pop))
df_long_pop <- cbind(df_long_pop, which_df)
which_df <- rep("pop0", nrow(df_long_pop_only))
df_long_pop_only <- cbind(df_long_pop_only, which_df)

# merge df together
df_long_both <- rbind(df_long, df_long_pop)
df_long_red <- rbind(df_long_pop, df_long_pop_only)
```


The null hypothesis of the Kolmogorov-Smirnov test is that the two samples come from the same distribution. In this following table, we compare the 'sample' where the population = 0 vs the rest. We see that the p-values are very small for every amenity, thus leading us to conclude that we have sufficient evidence to say that these 'samples' don't come from the same distribution. (We can conclude that there is an effect on the proximity measures when the population is 0 ?)

```{r}
amen_pval <-c()
n = 0
for (amen in amen_cols_short){
  n = n+1
  amen_pval[n]<- ks.test(filter(df_long_pop_only, df_long_pop_only$name == amen)$value, filter(df_long_pop, df_long_pop$name == amen)$value)[["p.value"]]
}

amen_pvaldf_red <- cbind(amen_cols_short,amen_pval)
amen_pvaldf_red
```
But in what ways do these subsets differ?

Here we are comparing the histogram for both, where the pink represents the count of population = 0, and the blue the rest. We see that the 'pink' appears to mirror the trends of the 'blue', but on a smaller scale. In the appendix, a 'zoomed in' plot is available. Surprisingly, see there that for some higher proximity 'bins' in transit and health, there are more cases for when population = 0. We also see in the appendix the kernel densities. 


```{r fig.height=10, fig.width=6}
# histograms
ggplot(df_long_red, aes(value, fill = which_df)) + 
    geom_histogram(bins = 50, alpha = .5, position="identity") + 
    facet_wrap(~name, scales = 'free_y', ncol = 2)
```

From this following table, we see that 72% of the proximity measure values where population = 0 are NA, compared to 50% of those where population !=0. 

```{r}
table(df_long_red$which_df,is.na(df_long_red$value))/cbind(table(df_long_red$which_df), table(df_long_red$which_df))
```
## Conclusion

There are no obvious clusters in the proximity measures to the naked eye.

The distribution of missing values is not the same across amenities nor provinces. 

Our dataset contains a lot of rows, and there may be question about the 'usefulness' of all of them. If we were to remove all the DBs where the population is 0, we could reduce our dataset by 23%, aiding computationally. There are still proximity measures associated with these DBs with population 0: the distributions of their proximity measures are not the statistically the same as those for the rest of the DBs (those with populations), but the trends appear somewhat similar. 

It doesn't appear that population of a DB is the only factor affecting whether a proximity measure is missing (NA); it was the only one tested as it was the only one included in this dataset. According to StatsCan's definition of a DB however, "only population and dwelling counts are disseminated at the dissemination block level" anyways. If we wanted to analyse other factors, we would have to look into aggregation at a higher level, which is not straighforward (need to take the mean/etc of whether something is missing or not?). 

 

\clearpage

# Appendix

```{r}
ggplot(pmd_na, aes(DBPOP)) + 
    geom_density()+labs(title = "Kernel density of DB population")
```

```{r fig.height=10, fig.width=12}
# Missing values for amennity by prov
# flattening so can facet
df_perc_missing %>% pivot_longer(all_of(amen_cols)) %>%
  ggplot(aes(y = value, x = name, fill = name)) + 
    geom_col() + 
    facet_wrap(~prov_name, ncol = 2) + labs(title = "Missing values for each amenity by province")
```

```{r}
# how many DBs by province
ggplot(pmd_na, aes(pr_abbr)) + geom_bar() + labs(title = "How many DBs by province")
```
## Logistics curves NAs

```{r fig.height=3, fig.width=3}
ggplot(df_na, aes(x=DBPOP, y=prox_idx_pharma)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_childcare)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_health)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_grocery)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_educpri)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_educsec)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_lib)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_parks)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
ggplot(df_na, aes(x=DBPOP, y=prox_idx_transit)) + 
  geom_point(alpha=.5) +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial))
```

## Histograms NAs

```{r fig.height=4, fig.width=6}
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_pharma)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_childcare)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_health)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_grocery)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_educpri)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_educsec)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_lib)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_parks)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
ggplot(df_na) +
  geom_histogram(aes(x=DBPOP, fill=factor(prox_idx_transit)), bins = 40,
                 colour="grey50", alpha=0.5, position="identity")
```

## Proximity measures histograms 

```{r fig.height=10, fig.width=6}
# histograms
# Next we see the histograms of proximity values for each amenity. This gives us an idea of the counts for each bin of values. We see that libraries have the least amount of proximity values that are near 0, but we saw above that they also had the greatest amount of missing values.
ggplot(df_long, aes(value)) + 
    geom_histogram(bins = 50) + 
    facet_wrap(~name, scales = 'free_y', ncol = 2)
```

## Proximity measures density zoomed in ? I think something is off
```{r fig.height=10, fig.width=6}
# kernel densities
ggplot(df_long, aes(value)) + 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 2) + xlim(0.5,1)+ labs(title = "Distribution of proximity measures by amenity, zoomed in")
```
## Population 0

```{r fig.height= 4, fig.width=6, include = F}
# violin plot
ggplot(df_long_pop, aes(x=name, y = value)) + geom_violin()
```

```{r fig.height=10, fig.width=6}
# histograms
ggplot(df_long_red, aes(value, fill = which_df)) + 
    geom_histogram(bins = 50, alpha = .5, position="identity") + 
    facet_wrap(~name, scales = 'free_y', ncol = 2) +xlim(0.25,1) + labs(title = 'Histogram of Population = 0 vs != 0, zoomed in')
```


```{r fig.height=10, fig.width=6}
# histograms
ggplot(df_long_red, aes(value, fill = which_df)) + 
    geom_density( alpha = .5, position="identity") + 
    facet_wrap(~name, scales = 'free_y', ncol = 2) +xlim(0.25,1) + labs(title = 'Kernel density of Population = 0 vs != 0, zoomed in')
```

## Table of counts for populations and NAs

There are still many more DBs that have populations > 0:

```{r}
table(df_long_red$which_df,is.na(df_long_red$value))
```
