---
title: "Personal Logs - Ricky"
author: 'Ricky Heinrich'
date: "MDS capstone"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(dplyr)
```

The following is a recommended way of keeping track of your weekly hours using this RMarkdown file.
Note that it will do all the calculations for you, you just need to enter in hours worked and a description of the task in the correct spot.

```{r}

```

## Instructions

1. For each week, first set the `current.week` to the week number (in most cases, you can increment the `current.week` by 1).\

2. Add row entries to your table indicating the date of the task, the number of hours spent on the task, a description of the task, as well any comments (or reminders to yourself) or requests for follow-up.

3. Knit the file to a markdown file.

4. Before Friday at 5 PM, please submit a Pull Request to merge in your weekly log. All your team mates will review your log, provide comments and suggestions, and then approve (or request changes). Once all the changes have been resolved, you can merge in your pull request (by Monday at 10 AM).

5. Don't forget to review the Pull Requests by your team mates!

6. Discuss any discrepancies or team dynamics issues with your assigned TA during the weekly meetings.

**Remember that your `markdown` file is what will be reviewed by your teammates, the TAs, and the instructors, so just make sure it renders correctly!**

## Week 0

```{r}
current.week = 1


mylog <- data.frame(week = current.week,
                       day=as.Date("2023-04-14"), 
                       hours=1,
                       description="Client Meeting",
                       comments="First meeting"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-04-21"), 
                       hours=0.5,
                       description="Client Meeting",
                       comments="Second meeting"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-04-28"), 
                       hours=2,
                       description="Client Meeting + presentation",
                       comments="Third meeting, guest presentation"
                       )

kable(mylog)
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?



1. What was your most significant **team contribution** to the project this week?

1. Are there any major blockers or roadblocks preventing you from proceeding on your project?

No, all ready to start (assuming no problems arised from instructor meeting)

## Week 1

Add individual contributions here (be sure to call the **tidyverse** and **knitr** package first)
```{r}
# current.week = current.week + 1

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=2.5,
                       description="capstone kickoff lecture",
                       comments="NA"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=2,
                       description="Code of Conduct and Team Contract",
                       comments="Group work"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=1,
                       description="Reading, prep",
                       comments="Read literature summaries, prepare for meeting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=1,
                       description="prep, Daily group meeting",
                       comments="Discuss tasks for day")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=0.5,
                       description="Plagiarism module",
                       comments="NA")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=0.5,
                       description="How to run a meeting video",
                       comments="NA")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=3,
                       description="Literature review",
                       comments="Searching, reading, citing")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-04"), 
                       hours=0.5,
                       description="Daily group meeting",
                       comments="work update")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-04"), 
                       hours=5,
                       description="Proposal work",
                       comments="Group work")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=1.5,
                       description="Client meeting",
                       comments="prep + feedback on draft proposal")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=2,
                       description="Proposal + slides",
                       comments="Edit proposal, create monday meeting slides. Group work")


kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

We did a lot of group work this week: creating the team contract and code of conduct, working on the proposal, and working on the slides for the monday meeting. We split off into two groups to draft the proposal report, half of us doing the literature review and proposal introduction, and the other half drafting the methods, research questions, and schedule. Through my literature review, I found out about the fuzzy c-means clustering algorithm and added it to the list to try out. We all came together to put everything together and review and elevate each other's work. There are no major roadblocks preventing us from proceeding in our project.



## Week 2

```{r}
current.week = current.week + 1
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-08"), 
                       hours=1.5,
                       description="Meetings",
                       comments="Group meeting (prep, daily), instructor meeting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-08"), 
                       hours=4.5,
                       description="EDA",
                       comments="shape files, windows, crs, exploring missing values"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-09"), 
                       hours=1,
                       description="Meetings",
                       comments="TA meeting, team meeting"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-09"), 
                       hours=2,
                       description="Team bonding",
                       comments="Connecting with team members for good dynamics"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-09"), 
                       hours=4.5,
                       description="EDA",
                       comments="attempt ppp, missing data analysis"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-10"), 
                       hours=0.5,
                       description="Daily Meeting",
                       comments="Updates, game plan"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-10"), 
                       hours=5.5,
                       description="EDA",
                       comments="distributions, summaries, missing values, draft"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-11"), 
                       hours=3,
                       description="Daily Meeting, EDA",
                       comments="Meeting + EDA review & discussion"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-11"), 
                       hours=0.5,
                       description="Proposal",
                       comments="Review proposal feedback"
                       )
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-11"), 
                       hours=4,
                       description="EDA",
                       comments="Pop = 0, etc, clean report"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-12"), 
                       hours=4.5,
                       description="Proposal",
                       comments="group work"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-12"), 
                       hours=1,
                       description="Misc",
                       comments="logs + misc over the week"
                       )

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

My most significant individual contribution to the project this week was the work I put into the EDA, and looking into what was the distribution of proximity measures when the population of a DB is 0 vs when it is not. I found out about and applied the Kolmogorov-Smirnov test. The most significant team contribution was talking with Jesse, who sent us resources about packages that help deal with NA values, to get us on the right track. The team started to investigate and apply these resources. This week we had major blockers of memory issues and computational power. We've somewhat solved the memory issues for some of us, but lack of computational power leads to waiting a long time for something to run, which doesn't feel efficient.  


## Week 3

```{r}
current.week = current.week + 1
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-15"), 
                       hours=3.5,
                       description="Meetings",
                       comments="Group meeting (prep, daily), instructor meeting, client meeting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-15"), 
                       hours=2,
                       description="Proposal",
                       comments="final edits")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-15"), 
                       hours=2.5,
                       description="EDA",
                       comments="outlier distribution, misc exploration")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-16"), 
                       hours=2,
                       description="Meetings",
                       comments="Group meeting, Data makers meeting, TA meeting, ")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-16"), 
                       hours=1,
                       description="EDA",
                       comments="misc exploration, team member's code")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-16"), 
                       hours=1,
                       description="Resource reading",
                       comments="categories of remoteness index")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-16"), 
                       hours=1,
                       description="Managing",
                       comments="Prep midterm presentation, minutes/agenda doc")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-17"), 
                       hours=1.5,
                       description="Meetings",
                       comments="Group meeting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-17"), 
                       hours=1.5,
                       description="EDA feedback",
                       comments="Reading and giving feedback on EDA report")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-18"), 
                       hours=2,
                       description="Misc",
                       comments="Reading proximity report, meeting, prep")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-19"), 
                       hours=1,
                       description="Meetings",
                       comments="Group meeting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-19"), 
                       hours=1,
                       description="Misc",
                       comments="Logs, reviewing, etc")



kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

My most significant individual contribution was reminding my team members that scaling is important and will change results when running some clustering algorithms, like the k-means. Our most significant group contribution is the group meeting with the DEIL team that made the PMD dataset: it gave us a lot of insight into the NA values and interpretation of proximity values, as well as insight into DB and different geographical areas. Some roadblocks this week include long running times, interpretation issues, and incongruence in feedback and acceptance of it. This week was short on hours because I was traveling from Thursday on to my Brother's grad out of province.

## Week 4

```{r}
current.week = current.week + 1
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-20"), 
                       hours=1,
                       description="Communication reflection",
                       comments="NA")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-21"), 
                       hours=1,
                       description="Midterm pres",
                       comments="design drafting")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-22"), 
                       hours=1,
                       description="Midterm pres",
                       comments="Outline")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-23"), 
                       hours=.5,
                       description="Agenda",
                       comments="meeting prep")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-23"), 
                       hours=3,
                       description="Meetings",
                       comments="Daily group, TA, client")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-23"), 
                       hours=1,
                       description="Midterm pres",
                       comments="reading report, visualizations, slides")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-24"), 
                       hours=1,
                       description="Meeting",
                       comments="daily meeting + agenda prep")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-24"), 
                       hours=6,
                       description="Midterm pres",
                       comments="work together slides, flow, practice pres")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-25"), 
                       hours=4.5,
                       description="class presentations",
                       comments="attending presentations")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-25"), 
                       hours=.5,
                       description="Prep",
                       comments="agenda prep + personal reflection")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-25"), 
                       hours=2,
                       description="Team meeting",
                       comments="presentation reflection, agendas, timeline")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-26"), 
                       hours=2,
                       description="Meetings",
                       comments="Jesse meeting, client meeting, team meeting")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-26"), 
                       hours=1.5,
                       description="Misc managing",
                       comments="schedule timeline, reformat agendas/notes docs")
mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-26"), 
                       hours=1,
                       description="Misc week",
                       comments="logs, prep")

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

My most significant contribution was implementing agendas for all of our meetings, including preparing them and moderating meetings. This has helped a lot for accountability and to use the meeting time meaningfully. Our most significant group contribution was making and presenting the midterm presentation. It helped center and reflect on our project. Roadblocks this week include unexpected results, showing the need to investigate algorithms and validation techniques some more. 

## Summary Table

To get the total for all the weekly hourly summary over each week use the following (requires the **dplyr** package)
```{r}
mylog %>% 
  group_by(week) %>% 
  summarise(total_weekly_hours = sum(hours)) %>%
  kable()
```
