---
title: "Personal Logs - Noman"
author: 'Noman Mohammad'
date: "MDS capstone"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(dplyr)
```

The following is a recommended way of keeping track of your weekly hours using this RMarkdown file.
Note that it will do all the calculations for you, you just need to enter in hours worked and a description of the task in the correct spot.

```{r}
mylog <- data.frame(week = 1, 
                    day = as.Date("2023-05-02"),
                    hours = 2.5,
                    description = "capstone kickoff lecture",
                    comments = NA
                    )

```

## Instructions

1. For each week, first set the `current.week` to the week number (in most cases, you can increment the `current.week` by 1).\

2. Add row entries to your table indicating the date of the task, the number of hours spent on the task, a description of the task, as well any comments (or reminders to yourself) or requests for follow-up.

3. Knit the file to a markdown file.

4. Before Friday at 5 PM, please submit a Pull Request to merge in your weekly log. All your team mates will review your log, provide comments and suggestions, and then approve (or request changes). Once all the changes have been resolved, you can merge in your pull request (by Monday at 10 AM).

5. Don't forget to review the Pull Requests by your team mates!

6. Discuss any discrepancies or team dynamics issues with your assigned TA during the weekly meetings.

**Remember that your `markdown` file is what will be reviewed by your teammates, the TAs, and the instructors, so just make sure it renders correctly!**

## Week 1

Add individual contributions here (be sure to call the **tidyverse** and **knitr** package first)
```{r}
current.week = 0
current.week = current.week + 1

mylog <- data.frame(week = 1, 
                    day = as.Date("2023-05-01"),
                    hours = 3,
                    description = "Prior meetings with client + presenation ",
                    comments = "client organized a presentation for our group"
                    )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=3,
                       description="Capstone kick off lecture",
                       comments=NA
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=2,
                       description="Team meeting + repo documents",
                       comments="after kickoff lecture, met with group and worked on documents for repo"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=7,
                       description="Team meeting + Working on proposal",
                       comments="plan work for proposal in meeting, read about project and dataset etc."
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-04"), 
                       hours=6,
                       description="Team meeting + Working on proposal",
                       comments="consolidate work on proposal, finaize before sending to client"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=.45,
                       description="Meeting with client",
                       comments="proposal feedback + follow up questions"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=2,
                       description="Team meeting + reforming proposal + slides",
                       comments="reform proposal, work on team-meeting slides for week1"
                       )

kable(mylog)
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?
feeling surprised that I made it to the capstone. Working on parts of the second half of the proposal.

1. What was your most significant **team contribution** to the project this week?
working on the proposal and meeting slides.

1. Are there any major blockers or roadblocks preventing you from proceeding on your project?
Nah

```{r}
current.week = current.week + 1

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-08"),
                    hours = 1.5,
                    description = "Team meeting + instructor meeting",
                    comments = "Final touches/planning on instructor meeting slides + meeting with professors"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-08"),
                    hours = 4.5,
                    description = "Exploring datasets",
                    comments = "Looking at potential data sets to merge"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-09"),
                    hours = 1,
                    description = "TA meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-09"),
                    hours = 2,
                    description = "Plotting merged dataset with IR",
                    comments = "lots of memory issues..."
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-10"),
                    hours = 3,
                    description = "Researching models that can deal with NA's (sent by jessie)",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-10"),
                    hours = 1,
                    description = "Update proposal section to revamp new methods",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-10"),
                    hours = 3,
                    description = "Implement mixAll clustering methods ",
                    comments = "Doing various things that Jonah had formualted in clustering doc"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-11"),
                    hours = 3,
                    description = "Researching Amelia and ways to impute ",
                    comments = "Implement Amelia, debugging. This can allow for checking correlation and apply PCA before clustering"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-11"),
                    hours = 3,
                    description = "Trying dimensionality reduction with clustering",
                    comments = "Implement various things after imputing values for data... interpret results"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-12"),
                    hours = 3,
                    description = "Team meeitng + proposal doc",
                    comments = "Had to leave early for Friday prayer, briefly tuned back in as team wrapped up"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-12"),
                    hours = 1,
                    description = "Random",
                    comments = "Reviewing logs, githib things etc.."
                    )

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?
Researching ways to impute values and dynamic clustering methods.

2. What was your most significant **team contribution** to the project this week?
Working on the team proposal document. 

3. Are there any major blockers or roadblocks preventing you from proceeding on your project?
trying to figure out a way to deal with NA's for tasks such as PCA etc. Also running into memory issues and execution times...

```{r}
current.week = current.week + 1

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-15"),
                    hours = 1,
                    description = "Weekly team meeting slides + practice",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-15"),
                    hours = 3.5,
                    description = "Team + Dr.Moosvi + Client meeting",
                    comments = NA
                    )


mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-16"),
                    hours = 3.5,
                    description = "Team + statscan (DEIL) + TA meetings ",
                    comments = "The statscan meeting was super helpful as all the gaps in our understanding about the data set were answered"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-16"),
                    hours = 0.5,
                    description = "Fixing git issue",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-16"),
                    hours = 1,
                    description = "Reading suggested reading after meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-17"),
                    hours = 1,
                    description = "Team meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-17"),
                    hours = 2.5,
                    description = "EDA review/edits",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-18"),
                    hours = 0.5,
                    description = "Team meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-18"),
                    hours = 4,
                    description = "Applying mixall method to template Jonah made",
                    comments = "leaving computer run over night due to runtime"
                    )


mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-19"),
                    hours = 3,
                    description = "Updated Mixall results with log transorms and various model disitributions",
                    comments = "very long running time...."
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-19"),
                    hours = 2,
                    description = "reading papers ricky suggested while running algorithms",
                    comments = NA
                    )

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?
Running MixALL algo for clustering

2. What was your most significant **team contribution** to the project this week?
Meeting with the DEIL lab

3. Are there any major blockers or roadblocks preventing you from proceeding on your project?
Just running times, but we are utilizing subsampling which is helping. 

```{r}
current.week = current.week + 1

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-23"),
                    hours = 3,
                    description = "Meeting with Team + Client + TA",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-23"),
                    hours = 3,
                    description = "Computing cutoff values",
                    comments = "getting data rdy for plots for final prez"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-23"),
                    hours = 1,
                    description = "Brainstorming slides",
                    comments = "thinking of slide structure for prez for modeling section"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-24"),
                    hours = 7,
                    description = "team meeting + slides + presentation practice",
                    comments = "trying to debug/fix cutoff values for MixAll as they didnt seem right after looking at visuals.."
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-25"),
                    hours = 4.5,
                    description = " Midterm presentation",
                    comments = "10 - 2:30"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-25"),
                    hours = 2,
                    description = "Team meeting + reflection",
                    comments = "reflection on our presentation and next steps for coming weeks"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-26"),
                    hours = 1,
                    description = "Meeting with TA",
                    comments = "Q and A session with TA"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-26"),
                    hours = 1,
                    description = "Meeting with client",
                    comments = "Re did our presentation with TA and talked about next steps"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-26"),
                    hours = 2.5,
                    description = "Trying MixAll with different transformation and trying to debug what is happening with MixAll method",
                    comments = ""
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-26"),
                    hours = 1,
                    description = "random thing throughout week",
                    comments = "reviewing logs/having to fix stashed changes."
                    )


kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?
Finding cutoff values for the MixAll method.

2. What was your most significant **team contribution** to the project this week?
Contributing to our mideterm presentation.

3. Are there any major blockers or roadblocks preventing you from proceeding on your project?
not really this week.

```{r}
current.week = current.week + 1

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-29"),
                    hours = 1.5,
                    description = "Team + Mentor meeting/ making slides",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-29"),
                    hours = 3,
                    description = "added new metrics + Method for ranking number of clusters",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-29"),
                    hours = 1,
                    description = "Re running algos with new added metrics",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-29"),
                    hours = 1,
                    description = "Research",
                    comments = "methods robust to outliers... these are mostly all denisty based methods that can handle noise but for distance based, kmeans mediods (PAM) seems promising"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-30"),
                    hours = 1.5,
                    description = "Team meeting + TA",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-30"),
                    hours = 6,
                    description = "Rewrote cutoff function to test if working properly, Changed way for finding optimal cluster function with Jesse's suggestion, Mclust, k means with PAM (robust to outliers)",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-31"),
                    hours = 1,
                    description = "Team meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-05-31"),
                    hours = 4,
                    description = "DENCLUE model (density based)",
                    comments = "found denclue 2.0 paper. Implementation exists in R? through some paid course online. Attempted to fit to Python implementation. fail... "
                    )

  mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-06-01"),
                    hours = 2.5,
                    description = "Team meeting/dynamics discussion",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-06-01"),
                    hours = 3,
                    description = "algorithm assumptions and how they work for previosuly fitted models",
                    comments = "final report"
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-06-02"),
                    hours = 1,
                    description = "client meeting",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-06-02"),
                    hours = 2,
                    description = "cluster profiling",
                    comments = NA
                    )

mylog <- mylog %>% add_row(week = current.week, 
                    day = as.Date("2023-06-02"),
                    hours = 1,
                    description = "random",
                    comments = "logs/emailing Jeff/ git"
                    )

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?
Fitting Mclust model and also a k-means model called k-mediod, which is an algorithm that is robust to outliers. The reason being was to see whether we get better metrics scores if we directly handle these outliers. Surprisingly, the cutoff values were very similiar to MixAll for some ammenities.

2. What was your most significant **team contribution** to the project this week?
Participating in our team dynamics meeting. I felt it was super valuable to gauge how everyone in our group was feeling mentally. We had a great group discussion and ironed out any concerns thus far. 

3. Are there any major blockers or roadblocks preventing you from proceeding on your project?
I think its important that we can find an agreed upon way to find a best clustering algorithm with the results we have thus far so we can highlight the group profiles. 

## Summary Table

To get the total for all the weekly hourly summary over each week use the following (requires the **dplyr** package)
```{r}
mylog %>% 
  group_by(week) %>% 
  summarise(total_weekly_hours = sum(hours)) %>%
  kable()
```
