---
title: "Personal Logs - Avishek"
author: 'Avishek Saha'
date: "MDS capstone"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(dplyr)
```

<!-- 
The following is a recommended way of keeping track of your weekly hours using this RMarkdown file.
Note that it will do all the calculations for you, you just need to enter in hours worked and a description of the task in the correct spot.
-->

```{r eval=FALSE}
mylog <- data.frame(week = 1, 
                    day = as.Date("2023-05-02"),
                    hours = 2.5,
                    description = "capstone kickoff lecture",
                    comments = NA
                    )
```

<!--

## Instructions

1. For each week, first set the `current.week` to the week number (in most cases, you can increment the `current.week` by 1).\

2. Add row entries to your table indicating the date of the task, the number of hours spent on the task, a description of the task, as well any comments (or reminders to yourself) or requests for follow-up.

3. Knit the file to a markdown file.

4. Before Friday at 5 PM, please submit a Pull Request to merge in your weekly log. All your team mates will review your log, provide comments and suggestions, and then approve (or request changes). Once all the changes have been resolved, you can merge in your pull request (by Monday at 10 AM).

5. Don't forget to review the Pull Requests by your team mates!

6. Discuss any discrepancies or team dynamics issues with your assigned TA during the weekly meetings.

**Remember that your `markdown` file is what will be reviewed by your teammates, the TAs, and the instructors, so just make sure it renders correctly!**

## Week 0 (Sample Week)
-->

```{r eval=FALSE}
current.week = 0

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), # for today's date
                       hours=2,
                       description="Activity during class to figure this out",
                       comments="Refreshed memory on RMarkdown"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"),
                       hours=3,
                       description="Troubleshooted RStudio and my R installation",
                       comments=""
                       )


kable(mylog)
```
<!-- My total weekly hours are: `r # sum(mylog[which(mylog$week==current.week), 3])` -->

<!--
### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?

1. What was your most significant **team contribution** to the project this week?

1. Are there any major blockers or roadblocks preventing you from proceeding on your project?
-->

## Week 1

<!-- Add individual contributions here (be sure to call the **tidyverse** and **knitr** package first) -->
```{r}
current.week = 1

mylog <- data.frame(week = current.week, 
                    day = as.Date("2023-05-01"),
                    hours = 3,
                    description="Initial Client Meetings",
                    comments="Meetings prior to May 1 (2 meetings + 1 presentation)"
                    )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-01"), 
                       hours=2,
                       description="Project Inception",
                       comments="Read & Understand Project Details"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-01"), 
                       hours=2,
                       description="Project Management",
                       comments="Create Notion for organizing all project work and timeline in one place"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=2.5,
                       description="Capstone Kickoff Lecture",
                       comments=NA
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=1.5,
                       description="Team Meeting",
                       comments="Discuss workflow"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-02"), 
                       hours=1.5,
                       description="Literature Review",
                       comments="Read peer reviewed research papers"
                       )


mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=4,
                       description="Literature Review",
                       comments="Read peer reviewed research papers"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=1,
                       description="Plagarism Module",
                       comments=NA
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-03"), 
                       hours=1,
                       description="Proposal Writing",
                       comments="Write a draft of the introduction"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-04"), 
                       hours=6,
                       description="Proposal",
                       comments="Team meeting on merge proposal"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=1,
                       description="Client Meeting",
                       comments="Feedback on Proposal"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=1.5,
                       description="Team Meeting",
                       comments="Finalize proposal and prepare slide decks for weekly meeting"
                       )

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-05"), 
                       hours=1,
                       description="Miscellaneous",
                       comments="Review Pull Requests, Store Logs, Organizing Notion etc."
                       )

# I don't have any comments to me

kable(mylog)
```
My total weekly hours are: `r sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

<!-- In this section, please write 5-7 sentences (in total) answering the following questions: -->

<!-- 1. What was your most significant **individual contribution** to the project this week? -->
- My most significant individual contribution to the project this week was likely the literature review, conducted on clustering proximity measures and how to evaluate different clustering techniques.

<!-- 2. What was your most significant **team contribution** to the project this week? -->
- This week was likely working on the proposal with team members. Specifically, splitting off into two groups to draft different sections of the proposal, then coming back together to review each other's work. This collaboration and teamwork helped move the project forward and ensured that everyone was on the same page.

<!-- 3. Are there any major blockers or roadblocks preventing you from proceeding on your project? -->
- There were no major blockers or roadblocks preventing me/us from proceeding on our project.

<!--
## Week 2

Add individual contributions here (be sure to call the **tidyverse** and **knitr** package first)
```{r eval=FALSE}
current.week = current.week + 1
mylog <- mylog %>% add_row(week = current.week,
                       day=Sys.Date(), # for today's date
                       hours=0.5,
                       description="learning this personal log rmd file",
                       comments="Rmd is cool")

mylog <- mylog %>% add_row(week = current.week,
                       day=as.Date("2023-05-01"), 
                       hours=1.5,
                       description="Editing kickoff slide decks with Firas",
                       comments="Firas will push the changes to repo"
                       )

kable(filter(mylog, week == current.week))
```
My total weekly hours are: `r # sum(mylog[which(mylog$week==current.week), 3])`

### Weekly Reflection

In this section, please write 5-7 sentences (in total) answering the following questions:

1. What was your most significant **individual contribution** to the project this week?

1. What was your most significant **team contribution** to the project this week?

1. Are there any major blockers or roadblocks preventing you from proceeding on your project?

## Summary Table

To get the total for all the weekly hourly summary over each week use the following (requires the **dplyr** package)
-->
```{r}
mylog %>% 
  group_by(week) %>% 
  summarise(total_weekly_hours = sum(hours)) %>%
  kable()
```