---
title: "Optics"
author: "Avishek"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=TRUE, 
                      fig.align="center", 
                      fig.width=8, 
                      fig.height=6)
```


* * *

## Introduction

OPTICS stands for Ordering Points To Identify Clustering Structure.  
This algorithm can be seen as a generalization of DBSCAN. A major issue with DBSCAN is that it fails to find clusters of varying density due to fixed eps.  
This is solved in OPTICS by using an approach of finding reachability of each point from the core points and then deciding the clusters based on reachability plot.  


Relevant terminologies for OPTICS

- $\epsilon$, epsilon (eps): is known as the Maximum allowed distance from one point to the other point for both of them to be considered in one group/cluster
- MinPts: is the minimum number of points which should be present close to each other at a distance of epsilon ($\epsilon$) so that they all can be form a group/cluster
Core Point: That point which has at least MinPts number of points near to it, within the distance of $\epsilon$(eps)
- Border Point/Non-Core Point: That point in data which has less than the minimum number of points(MinPts) within its reach (a distance of eps)
- Noise: That point which has no point near to it within a distance of eps
- Core Distance: The minimum distance required by a point to become a core point. It means it is possible to find the MinPts number of points within this distance. Core distance can be less than the pre-decided value of $\epsilon$, epsilon (eps), which is the maximum allowed distance to find MinPts.
- Reachability distance: This is the distance to reach a point from the cluster. Now if the point lies within the Core Distance, then Reachability Distance=Core Distance. And, if the point lies outside the Core Distance, then Reachability Distance is the minimum distance from the extreme point of the cluster.

## Alogrithm Steps
1. For the given values of MinPts and eps($\epsilon$). Find out if a point is close to MinPts number of points within a distance less than or equal to eps. Tag it as a Core Point. Update the reachability distance = core distance for all the points within the cluster.
2. If it is not a core point then find out its density connected distance from the nearest cluster. Update the reachability distance.
3. Arrange the data in increasing order of reachability distance for each cluster. The smallest distances come first and represent the dense sections of data and the largest distances come next representing the noise section. This is a special type of dendrogram
4. Find out the places where a sharp decline is happening in the reachability distance plot.
5. “Cut” the plot in the y-axis by a suitable distance to get the clusters.


## Loading \& Cleaning Data

```{r}
set.seed(2023)
library(dplyr)
library(cluster)
library(ggplot2)
library(factoextra)
library(clusterCrit)
library(clustertend)
load('../../../data/master_pms_df.Rdata')

# Convert selected columns to numeric
master <- master %>%
  mutate(PMS_DBPOP = as.numeric(gsub(",", "", PMS_DBPOP)), # Dissemination block population
         PMS_DAPOP = as.numeric(gsub(",", "", PMS_DAPOP)), # Dissemination area population
         PMS_CSDPOP = as.numeric(gsub(",", "", PMS_CSDPOP)), # Census subdivision population
         PMS_CMAPOP = as.numeric(gsub(",", "", PMS_CMAPOP)), # Census metropolitan area population
         PMS_PRPOP = as.numeric(gsub(",", "", PMS_PRPOP)), # Province or territory population
         # in_db_emp = as.numeric(in_db_emp),
         # in_db_pharma = as.numeric(in_db_pharma),
         # in_db_childcare = as.numeric(in_db_childcare),
         # in_db_health = as.numeric(in_db_health),
         # in_db_grocery = as.numeric(in_db_grocery),
         # in_db_educpri = as.numeric(in_db_educpri),
         # in_db_educsec = as.numeric(in_db_educsec),
         # in_db_lib = as.numeric(in_db_lib),
         # in_db_parks = as.numeric(in_db_parks),
         # in_db_transit = as.numeric(in_db_transit),
         PMS_prox_idx_emp = as.numeric(PMS_prox_idx_emp),
         PMS_prox_idx_pharma = as.numeric(PMS_prox_idx_pharma),
         PMS_prox_idx_childcare = as.numeric(PMS_prox_idx_childcare),
         PMS_prox_idx_health = as.numeric(PMS_prox_idx_health),
         PMS_prox_idx_grocery = as.numeric(PMS_prox_idx_grocery),
         PMS_prox_idx_educpri = as.numeric(PMS_prox_idx_educpri),
         PMS_prox_idx_educsec = as.numeric(PMS_prox_idx_educsec),
         PMS_prox_idx_lib = as.numeric(PMS_prox_idx_lib),
         PMS_prox_idx_parks = as.numeric(PMS_prox_idx_parks),
         PMS_prox_idx_transit = as.numeric(PMS_prox_idx_transit),
         DBUID = as.character(DBUID),
         PMS_DAUID = as.character(PMS_DAUID),
         PMS_CSDUID = as.character(PMS_CSDUID),
         PMS_CMAUID = as.character(PMS_CMAUID),
         PMS_CMAPUID = as.character(PMS_CMAPUID),
         PMS_PRUID = as.character(PMS_PRUID),
         PMS_suppressed = as.character(PMS_suppressed),
         PMS_transit_na = as.character(PMS_suppressed))


# Subset columns that start with "prox_idx"
amenities <- colnames(master)[grepl("^PMS_prox_idx", colnames(master))]
# master dataset - contains only proximity columns
master_amenities <- master[, amenities]
# variables to cluster with
clust_vars = c('CSD_AREA', 'PMS_CSDPOP', 'PMS_DBPOP', 'IOR_Index_of_remoteness') #, 'PMS_CMATYPE')

# subsampling data 
perc = 3 #percentage of data to subsample
subsample = (nrow(master)/100)*perc 
master_sample = master[sample(nrow(master), subsample),]

master_sample_log <- master_sample

# log transform PMs
for (i in amenities){
  master_sample_log[,i] = log(master_sample[,i] + 0.0001)
}
```

```{r}
library(dbscan)
# Creating Clusters using OPTICS algorithm
Data <- as.matrix(na.omit(master_sample_log$PMS_prox_idx_emp))
OPTICSclusters=optics(Data, eps = 9, minPts = round(0.05 * length(Data)))
OPTICSclusters
```

```{r}
# Extract the Rechability Distances by cutting the reachability plot at eps_cl
# extractDBSCAN() extracts a clustering from an OPTICS ordering that is similar to what
# DBSCAN would produce with a single global ϵ set to eps_cl.
RechabilityDistances=extractDBSCAN(OPTICSclusters, eps_cl = 0.05)
# Plotting the reachability distances to see cluster formation
# Black lines are noise
plot(RechabilityDistances)
```

```{r}
unique(RechabilityDistances$cluster)
```


```{r}
# ordered datapoint
head(OPTICSclusters$order, n =15)
```


```{r}
plot(Data, col = "grey")
polygon(Data[OPTICSclusters$order,], )
```


```{r}
# it wont work in univariate case
# This function produces a two-dimensional scatter plot with added convex hulls for clusters.
hullplot(Data, OPTICSclusters)
```

extractXi() extracts clusters hierarchically based on the steepness of the reachability plot. minimum controls whether only the minimal (non-overlapping) cluster are extracted. correctPredecessor corrects a common artifact known of the original ξ method presented in Ankerst et al. (1999) by pruning the steep up area for points that have predecessors not in the cluster (see Technical Note in Appendix A for details).
extractXi(object, xi, minimum = FALSE, correctPredecessor = TRUE)

```{r}
# extract Clusters of varying density using the Xi method
XiClusters=extractXi(OPTICSclusters, xi = 0.03)
XiClusters
```

```{r}
XiClusters$clusters_xi
```


```{r}
dend <- as.dendrogram(XiClusters)
dend
```

```{r}
plot(dend, ylab = "Reachability dist.", leaflab = "none")
```


```{r}
ClusterData <- as.data.frame(Data)
# Assigning cluster id for each row
ClusterData$ClusterID=XiClusters$cluster
colnames(ClusterData) <- c("Data", "Cluster")
head(ClusterData)
```

```{r}
length(ClusterData[ClusterData$Cluster==5, "Cluster"])
```

```{r}
ClusterData[ClusterData$Data == -3.750755, "Cluster"]
```
```{r}
# Count the number of rows with data value -3.750755
numRows <- sum(ClusterData$Data == -6.725434)

# Print the result
print(numRows)

```


```{r}
cutoffs <- tapply(ClusterData$Data, ClusterData$Cluster, quantile, probs = c(0, 1))
```

```{r}
cutoffs
```


```{r}
# cut-off viz
# data frame with the cutoff values and group information
cutoff_data <- data.frame(Group = rownames(cutoffs))
cutoff_data$Start <- sapply(cutoffs, function(x) x["0%"])
cutoff_data$End <- sapply(cutoffs, function(x) x["100%"])

# vector of colors for each group
# cluster_colors <- c("1" = "red", "2" = "blue", "3" = "green", "4" = "purple", "5" = "orange")


plot(ggplot(cutoff_data) +
  geom_rect(aes(xmin = Start, xmax = End, ymin = 0, ymax = 0.1, fill = Group, alpha = 0.5)) +
  # scale_fill_manual(values = cluster_colors) +
  labs(x = "Emp", y = NULL,
       title = "Cutoff Values by Quartile") +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_cartesian(xlim = range(cutoff_data$Start, cutoff_data$End)))
```


```{r}
sill_coef <- intCriteria(as.matrix(ClusterData), 
                           as.integer(XiClusters$cluster), 
                                     'Silhouette')$silhouette
sill_coef
```

```{r}
# bounded between 0 and 1. Lower is better.
library(clusterSim)
davies_bouldin <- index.DB(ClusterData, XiClusters$cluster)$DB
davies_bouldin
```

## Grocery

```{r}
# Creating Clusters using OPTICS algorithm
Data_gro <- as.matrix(na.omit(master_sample_log$PMS_prox_idx_grocery))
OPTICSclusters_gro=optics(Data_gro, eps = 9, minPts = round(0.05 * length(Data_gro)))
OPTICSclusters_gro
```

```{r}
# extract Clusters of varying density using the Xi method
XiClusters_gro=extractXi(OPTICSclusters_gro, xi = 0.025)
XiClusters_gro
```

```{r}
ClusterData_gro <- as.data.frame(Data_gro)
# Assigning cluster id for each row
ClusterData_gro$ClusterID=XiClusters_gro$cluster
colnames(ClusterData_gro) <- c("Data", "Cluster")
head(ClusterData_gro)
```

```{r}
cutoffs_gro <- tapply(ClusterData_gro$Data, ClusterData_gro$Cluster, quantile, probs = c(0, 1))
```

```{r}
cutoffs_gro
```

```{r}
sill_coef_gro <- intCriteria(as.matrix(ClusterData_gro), 
                           as.integer(XiClusters_gro$cluster), 
                                     'Silhouette')$silhouette
```


```{r}
davies_bouldin_gro <- index.DB(ClusterData_gro, XiClusters_gro$cluster)$DB
davies_bouldin_gro
```


```{r}
# Creating Clusters using OPTICS algorithm
Data_h <- as.matrix(na.omit(master_sample_log$PMS_prox_idx_health))
OPTICSclusters_h=optics(Data_h, eps = 9, minPts = round(0.05 * length(Data_h)))
OPTICSclusters_h
```

```{r}
# extract Clusters of varying density using the Xi method
XiClusters_h=extractXi(OPTICSclusters_h, xi = 0.03)
XiClusters_h
```

```{r}
ClusterData_h <- as.data.frame(Data_h)
# Assigning cluster id for each row
ClusterData_h$ClusterID=XiClusters_h$cluster
colnames(ClusterData_h) <- c("Data", "Cluster")
head(ClusterData_h)
```

```{r}
cutoffs_h <- tapply(ClusterData_h$Data, ClusterData_h$Cluster, quantile, probs = c(0, 1))
```

```{r}
cutoffs_h
```

```{r}
sill_coef_h <- intCriteria(as.matrix(ClusterData_h), 
                           as.integer(XiClusters_h$cluster), 
                                     'Silhouette')$silhouette
```





































