---
title: "Clustering Template"
author: "PMS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=6, fig.align = 'center')
```


* * *

# Preliminary 

## Loading \& Cleaning Data

```{r, message=FALSE}
# loading libraries
set.seed(2023)
library(cluster)
library(ggplot2)
library(factoextra)
library(clusterCrit)

# loading data
# ...

# subsampling data (if needed)
# perc = 20 #percentage of data to subsample
# subsample = (nrow(master)/100)*perc 
# subsam = master[sample(nrow(master), subsample), idx]

# variables to cluster with
clust_vars = c('CSDTYPE_grouped, PMS_DBPOP, IOR_Index_of_remoteness')
```



## Assumptions of the Alogrithm

text



*** 


\pagebreak
# Employment 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_emp),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 


\pagebreak
# Pharmacy 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_pharma),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 


\pagebreak
# Childcare  

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_childcare),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 


\pagebreak
# Healthcare  

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_health),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 


\pagebreak
# Grocery 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_grocery),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 



\pagebreak
# Primary Education

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_educpri),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 



\pagebreak
# Secondary Education  

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_educsec),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 


\pagebreak
# Library 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_lib),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 



\pagebreak
# Parks 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_parks),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 



\pagebreak
# Transit 

## Implementation 

```{r}
# remove NA values
# amen = master[!is.na(master$PMS_prox_idx_transit),]

# algorithm
#

# plot
#pass = list(data = <cluster_data>, cluster = <cluster_assignments>)
#fviz_cluster(pass, ellipse.type = "norm") + theme_minimal()
```


## Cut-off Values

```{r}
# need to code this still
```

## Silhouette Plot

```{r}
# sil = silhouette(<cluster_assignments>, dist(<cluster_data>))
# fviz_silhouette(sil)
```


## Cluster Profiles

```{r}
# for (k in sort(unique(<cluster_assignments>))){
#   temp = amen[<cluster_assignments> == k,]
#   print(paste('Cluster #', k))
#   print(paste('Num of DBs in cluster: ', as.character(nrow(temp))))
#   print('CSD Type:')
#   print(table(temp$CSDTYPE_grouped)) 
#   cat('\n DB Population: \n')
#   print(summary(temp$PMS_DBPOP))
#   cat('\n Index of Remoteness: \n')
#   print(summary(temp$IOR_Index_of_remoteness))
#   cat('\n Provinces: \n')
#   print(table(temp$PROVINCE))
#   cat('\n Amenity dense: \n')
#   print(table(temp$PMS_amenity_dense))
#   cat('\n\n\n ')
# }

# save memory
# rm(amen)
```

*** 




\pagebreak
# Conclusion

text







