---
title: "Clustering Template"
author: "PMS"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=6, fig.align = 'center')
```


* * *

# Preliminary 

## Loading \& Cleaning Data

```{r}
# 
```



## Assumptions of the Alogrithm

text


*** 

*** 



# Individual Metrics

## Employment 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 


## Pharmacy 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 


## Childcare  

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 


## Healthcare  

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 


## Grocery 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 



## Primary Education

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 



## Secondary Education  

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 


## Library 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 



## Parks 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 



## Transit 

### Implementation 

```{r}
#
```


### Cut-off Values

```{r}
#
```

### Silhouette Plot

```{r}
#
```


### Cluster Profiles

```{r}
#
```

*** 




## Conclusion

text



*** 

*** 

\pagebreak
# All Metrics Together 

## Implementation 

```{r}
#
```


## Cut-off Values

```{r}
#
```

## Silhouette Plot

```{r}
#
```


## Cluster Profiles

```{r}
#
```

## Conclusion

text


***

***

\pagebreak
# Linked with Index of Remoteness


## Implementation 

```{r}
#
```


## Cut-off Values

```{r}
#
```

## Silhouette Plot

```{r}
#
```


## Cluster Profiles

```{r}
#
```

## Conclusion

text





