---
title: "JenksBreaks"
author: "Avishek"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=TRUE, 
                      warning = F,
                      message = F,
                      fig.align="center" 
                      )
```

* * *

## Introduction
The Jenks Natural Breaks Classification (or Optimization) system is a data classification method designed to optimize the arrangement of a set of values into "natural" classes.   
- A Natural class is the most optimal class range found "naturally" in a data set.   
- Natural breaks are determined with a frequency histogram. Class boundaries are identified as troughs in the data.
- Many dataset will not have obvious natural breaks which means that this method would tend to show breaks where none really exists.  

Jenks Natural Breaks Classification is done by seeking to minimize each class's average deviation from the class mean, while maximizing each class's deviation from the means of the other classes. In other words, the method seeks to reduce the variance within classes and maximize the variance between classes.

## Alogrithm Steps
Step 1: The user selects the attribute, x, to be classified and specifies the number of classes required, k

Step 2: A set of k‑1 random or uniform values are generated in the range [min{x},max{x}]. These are used as initial class boundaries

Step 3: The mean values for each initial class are computed and the sum of squared deviations of class members from the mean values is computed. The total sum of squared deviations (TSSD) is recorded

Step 4: Individual values in each class are then systematically assigned to adjacent classes by adjusting the class boundaries to see if the TSSD can be reduced. This is an iterative process, which ends when improvement in TSSD falls below a threshold level, i.e. when the within class variance is as small as possible and between class variance is as large as possible. True optimization is not assured. The entire process can be optionally repeated from Step 1 or 2 and TSSD values compared


```{r}
set.seed(2023)
library(dplyr)
library(cluster)
library(ggplot2)
library(factoextra)
library(clusterCrit)
library(clustertend)
library(fpc)
library(clusterSim)
library(BAMMtools)
load('../../../data/master_pms_df.Rdata')

# Convert selected columns to numeric
master <- master %>%
  mutate(PMS_DBPOP = as.numeric(gsub(",", "", PMS_DBPOP)), # Dissemination block population
         PMS_DAPOP = as.numeric(gsub(",", "", PMS_DAPOP)), # Dissemination area population
         PMS_CSDPOP = as.numeric(gsub(",", "", PMS_CSDPOP)), # Census subdivision population
         PMS_CMAPOP = as.numeric(gsub(",", "", PMS_CMAPOP)), # Census metropolitan area population
         PMS_PRPOP = as.numeric(gsub(",", "", PMS_PRPOP)), # Province or territory population
         # in_db_emp = as.numeric(in_db_emp),
         # in_db_pharma = as.numeric(in_db_pharma),
         # in_db_childcare = as.numeric(in_db_childcare),
         # in_db_health = as.numeric(in_db_health),
         # in_db_grocery = as.numeric(in_db_grocery),
         # in_db_educpri = as.numeric(in_db_educpri),
         # in_db_educsec = as.numeric(in_db_educsec),
         # in_db_lib = as.numeric(in_db_lib),
         # in_db_parks = as.numeric(in_db_parks),
         # in_db_transit = as.numeric(in_db_transit),
         PMS_prox_idx_emp = as.numeric(PMS_prox_idx_emp),
         PMS_prox_idx_pharma = as.numeric(PMS_prox_idx_pharma),
         PMS_prox_idx_childcare = as.numeric(PMS_prox_idx_childcare),
         PMS_prox_idx_health = as.numeric(PMS_prox_idx_health),
         PMS_prox_idx_grocery = as.numeric(PMS_prox_idx_grocery),
         PMS_prox_idx_educpri = as.numeric(PMS_prox_idx_educpri),
         PMS_prox_idx_educsec = as.numeric(PMS_prox_idx_educsec),
         PMS_prox_idx_lib = as.numeric(PMS_prox_idx_lib),
         PMS_prox_idx_parks = as.numeric(PMS_prox_idx_parks),
         PMS_prox_idx_transit = as.numeric(PMS_prox_idx_transit),
         DBUID = as.character(DBUID),
         PMS_DAUID = as.character(PMS_DAUID),
         PMS_CSDUID = as.character(PMS_CSDUID),
         PMS_CMAUID = as.character(PMS_CMAUID),
         PMS_CMAPUID = as.character(PMS_CMAPUID),
         PMS_PRUID = as.character(PMS_PRUID),
         PMS_suppressed = as.character(PMS_suppressed),
         PMS_transit_na = as.character(PMS_suppressed))


# Subset columns that start with "prox_idx"
amenities <- colnames(master)[grepl("^PMS_prox_idx", colnames(master))]
# master dataset - contains only proximity columns
master_amenities <- master[, amenities]
# variables to cluster with
clust_vars = c('CSD_AREA', 'PMS_CSDPOP', 'PMS_DBPOP', 'IOR_Index_of_remoteness') #, 'PMS_CMATYPE')

# subsampling data 
perc = 3 #percentage of data to subsample
subsample = (nrow(master)/100)*perc 
master_sample = master[sample(nrow(master), subsample),]

master_sample_log <- master_sample

# log transform PMs
for (i in amenities){
  master_sample_log[,i] = log(master_sample[,i] + 0.0001)
}

master_amenities_log <- master_amenities

for (i in amenities){
  master_amenities_log[,i] = log(master_amenities[,i] + 0.0001)
}
```



```{r}
# Mode function
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

algo <- function(data, c_num){
  avg_sil_width = c()
  dunn_stats = c()
  xie_beni = c()
  calinski_harabasz = c()
  davies_bouldin = c()
  
  counter = 1
  
  # run algo for different number of clusters
  for (i in c_num){
   
    # apply the jenks
    jenks_breaks <- getJenksBreaks(data$Data, i)
    # jenks_breaks_org <- exp(jenks_breaks) - 0.0001

  
    # max_val <- max(data$Data)
    
    # group numbers based on breaks
    data$Cluster <- NA

    for (i in 1:length(jenks_breaks)) {
      if (i == 1) {
        data$Cluster <- ifelse(data$Data <= jenks_breaks[i], i, NA)
      } else {
        data$Cluster <- ifelse(data$Data< jenks_breaks[i] & is.na(data$Cluster), 
                                          i, data$Cluster)
      }
    }
    
    # last group to values above the highest break
    data$Cluster <- ifelse(is.na(data$Cluster), 
                                      length(jenks_breaks) + 1, 
                                      data$Cluster)
    
    
    clusts <- data$Cluster
    comp_data <- data$Data
    
    sil = silhouette(clusts, dist(comp_data))
    
    avg_sil_width[counter] <- mean(sil[, "sil_width"])
    
    dunn_stats[counter] <- cluster.stats(dist(comp_data), clusts)$dunn
    
    xie_beni[counter] <- intCriteria(as.matrix(comp_data), 
                            as.integer(clusts), 
                            'Xie_Beni')$xie_beni
    
    calinski_harabasz[counter] <- calinhara(comp_data, clusts)
    
    davies_bouldin[counter] <- index.DB(comp_data, clusts)$DB
    
    counter = counter + 1
  }
  
  #plot silhouette coefficients
  plot(avg_sil_width~c_num, type = 'l', ylab="Silhouette Coefficient", xlab='Number of Clusters', lwd=2)

  # plot Dunn Index coefficients
  plot(dunn_stats~c_num, type = 'l', ylab="Dunn Coefficient", xlab='Number of Clusters', lwd=2)

  # plot Xie_Beni coefficients
  plot(xie_beni~c_num, type = 'l', ylab="Xie_Beni Coefficient", xlab='Number of Clusters', lwd=2)

  # plot Calinski Harabasz coefficients
  plot(calinski_harabasz~c_num, type = 'l', ylab="Calinski Harabasz Coefficient", xlab='Number of Clusters', lwd=2)

  # plot davies_bouldin coefficients
  plot(davies_bouldin~c_num, type = 'l', ylab="Davies Bouldin Coefficient", xlab='Number of Clusters', lwd=2)


  res_avg_sil_width <- c_num[which(avg_sil_width == max(avg_sil_width))]
  res_dunn_stats <- c_num[which(dunn_stats == max(dunn_stats))]
  res_xie_beni <- c_num[which(xie_beni == min(xie_beni))]
  res_calinski_harabasz <- c_num[which(calinski_harabasz == max(calinski_harabasz))]
  res_davies_bouldin <- c_num[which(davies_bouldin == min(davies_bouldin))]


  # the most suggested number of clusters
  most_frequent <- Mode(c(res_avg_sil_width,
                          res_dunn_stats,
                          res_xie_beni,
                          res_calinski_harabasz,
                          res_davies_bouldin))

  # re-run algorithm with most suggested number of clusters
  jenks_breaks <- getJenksBreaks(data$Data, most_frequent)

  return(list("jenks_breaks" = jenks_breaks, "no_clust" = most_frequent))
}

do_everything = function(data, num){
  
  data <- as.matrix(na.omit(data))
  # make a dataframe with clusters and data
  data <- as.data.frame(data)
  colnames(data) <- c("Data")

  res = algo(data, num)

  jenks_breaks <- res$jenks_breaks
  
  max_val <- max(data$Data)
  
# group numbers based on breaks
    data$Cluster <- NA

    for (i in 1:length(jenks_breaks)) {
      if (i == 1) {
        data$Cluster <- ifelse(data$Data <= jenks_breaks[i], i, NA)
      } else {
        data$Cluster <- ifelse(data$Data< jenks_breaks[i] & is.na(data$Cluster), 
                                          i, data$Cluster)
      }
    }
    
    # last group to values above the highest break
    data$Cluster <- ifelse(is.na(data$Cluster), 
                                      length(jenks_breaks) + 1, 
                                      data$Cluster)
  
  clusts <- data$Cluster
  comp_data <- data$Data
  
  sil = silhouette(clusts, dist(comp_data))
  plot(fviz_silhouette(sil))
  
  avg_sil_width <- mean(sil[, "sil_width"])
  
  dunn_stats <- cluster.stats(dist(comp_data), clusts)$dunn
  
  xie_beni <- intCriteria(as.matrix(comp_data), 
                          as.integer(clusts), 
                          'Xie_Beni')$xie_beni
  
  calinski_harabasz <- calinhara(comp_data, clusts)
  
  davies_bouldin <- index.DB(comp_data, clusts)$DB
  
  
  cutoffs <- tapply(data$Data, data$Cluster, quantile, probs = c(0, 1))

  # cut-off viz
  # data frame with the cutoff values and group information
  cutoff_data <- data.frame(Group = rownames(cutoffs))
  cutoff_data$Start <- sapply(cutoffs, function(x) x["0%"])
  cutoff_data$End <- sapply(cutoffs, function(x) x["100%"])

  return(list('avg_sill_width' = avg_sil_width,
              "dunn_index" = dunn_stats,
              "xie_beni" = xie_beni,
              "calinski_harabasz" = calinski_harabasz,
              "davies_bouldin" = davies_bouldin,
              "opt_n_clust" = res$no_clust,
              "cut_offs" = cutoff_data,
              "breaks" = jenks_breaks
              )
         )
}
```


```{r}
rug_plot <- function(cluster_data, Data){
  num_clusters <- length(unique(cluster_data$Group))
  cluster_colors <- hcl(h = seq(15, 375, length = num_clusters), c = 100, l = 65)
  
  ggplot(Data, aes(x = Value)) +
    geom_histogram(aes( y = ..density..),
                   colour = 1, fill = "white", bins = 50) +
    geom_density() +
    # rug cluster
    geom_segment(data = cluster_data, aes(x = Start, xend = End, 
                                          y = 0, yend = 0, 
                                          color = as.factor(Group)), 
                 size = 2) +
    # rug points
    geom_rug(data = Data, aes(x = Value), sides = "b", size = 0.2) +
    scale_color_manual(values = cluster_colors) +
    
    geom_vline(data = cluster_data, 
               aes(xintercept = Start), 
               color = "blue", linetype = "dashed", size = 0.1, alpha = 0.5) +
    
    geom_vline(data = cluster_data, 
               aes(xintercept = End), 
               color = "red", linetype = "dashed", size = 0.1, alpha = 0.5) +
    
    ylab("") +
    xlab("Value") +
    ggtitle("Cluster Rugs with Histogram & Cutoff Line") +
    theme_minimal() +
    coord_cartesian(xlim = c(min(cluster_data$Start) - 1, max(cluster_data$End) + 1))
}
```



### Employment 

```{r}
res_emp = do_everything(master_sample_log$PMS_prox_idx_emp, 2:10)
```

In the cutoff line figures below Blue vertical line indicates a cluster starting point and red vertical line indicates cluster's ending point. If you can't see the blue line before the red line that means either the cluster contains only one point or the starting and ending point of that cluster is so close. 

```{r}
Data_emp <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_emp))
rug_plot(res_emp$cut_offs, Data_emp)
```


```{r}
# org_cutoffs <- data.frame(
#   Cluster_no <- res_emp$cut_offs$Group,
#   Start = exp(res_emp$cut_offs$Start) - 0.0001,
#   End = exp(res_emp$cut_offs$End) - 0.0001
# )
```


### Health 

```{r}
res_health = do_everything(master_sample_log$PMS_prox_idx_health, 2:10)
```

```{r}
Data_health <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_health))
rug_plot(res_health$cut_offs, Data_health)
```


### Childcare 

```{r}
res_childcare = do_everything(master_sample_log$PMS_prox_idx_childcare, 2:10)
```

```{r}
Data_childcare <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_childcare))
rug_plot(res_childcare$cut_offs, Data_childcare)
```

### Parks 

```{r}
res_parks = do_everything(master_sample_log$PMS_prox_idx_parks, 2:10)
```


```{r}
Data_parks <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_parks))
rug_plot(res_parks$cut_offs, Data_parks)
```


### Education Primary 

```{r}
res_educpri = do_everything(master_sample_log$PMS_prox_idx_educpri, 2:10)
```

```{r}
Data_educpri <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_educpri))
rug_plot(res_educpri$cut_offs, Data_educpri)
```

### Transit 

```{r}
res_transit = do_everything(master_sample_log$PMS_prox_idx_transit, 2:10)
```


```{r}
Data_transit <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_transit))
rug_plot(res_transit$cut_offs, Data_transit)
```

### Pharma 

```{r}
res_pharma = do_everything(master_sample_log$PMS_prox_idx_pharma, 2:10)
```

```{r}
Data_pharma <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_pharma))
rug_plot(res_pharma$cut_offs, Data_pharma)
```

### Education Secondary 

```{r}
res_educsec = do_everything(master_sample_log$PMS_prox_idx_educsec, 2:10)
```

```{r}
Data_educsec <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_educsec))
rug_plot(res_educsec$cut_offs, Data_educsec)
```

### Grocery 

```{r}
res_grocery = do_everything(master_sample_log$PMS_prox_idx_grocery, 2:10)
```

```{r}
Data_grocery <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_grocery))
rug_plot(res_grocery$cut_offs, Data_grocery)
```

### Library 

```{r}
res_lib = do_everything(master_sample_log$PMS_prox_idx_lib, 2:10)
```


```{r}
Data_lib <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_lib))
rug_plot(res_lib$cut_offs, Data_lib)
```

```{r}
print("Silhouette Coefficients")
sill_quintiles <- c(res_emp$avg_sill_width, res_pharma$avg_sill_width, res_childcare$avg_sill_width, res_health$avg_sill_width, res_grocery$avg_sill_width, res_educpri$avg_sill_width, res_educsec$avg_sill_width, res_lib$avg_sill_width, res_parks$avg_sill_width, res_transit$avg_sill_width)
sill_quintiles
```

```{r}
print("Dunn Index")
dunn_quintiles <- c(res_emp$dunn_index, res_pharma$dunn_index, res_childcare$dunn_index, res_health$dunn_index, res_grocery$dunn_index, res_educpri$dunn_index, res_educsec$dunn_index, res_lib$dunn_index, res_parks$dunn_index, res_transit$dunn_index)
dunn_quintiles
```

```{r}
print("Xie Beni")
xie_beni_quintiles <- c(res_emp$xie_beni, res_pharma$xie_beni, res_childcare$xie_beni, res_health$xie_beni, res_grocery$xie_beni, res_educpri$xie_beni, res_educsec$xie_beni, res_lib$xie_beni, res_parks$xie_beni, res_transit$xie_beni)
xie_beni_quintiles
```

```{r}
print("Calinski Harabasz")
calinski_harabasz_quintiles <- c(res_emp$calinski_harabasz, res_pharma$calinski_harabasz, res_childcare$calinski_harabasz, res_health$calinski_harabasz, res_grocery$calinski_harabasz, res_educpri$calinski_harabasz, res_educsec$calinski_harabasz, res_lib$calinski_harabasz, res_parks$calinski_harabasz, res_transit$calinski_harabasz)
calinski_harabasz_quintiles
```

```{r}
print("Davies Bouldin")
davies_bouldin_quintiles <- c(res_emp$davies_bouldin, res_pharma$davies_bouldin, res_childcare$davies_bouldin, res_health$davies_bouldin, res_grocery$davies_bouldin, res_educpri$davies_bouldin, res_educsec$davies_bouldin, res_lib$davies_bouldin, res_parks$davies_bouldin, res_transit$davies_bouldin)
davies_bouldin_quintiles
```

```{r}
print("Number of clusters")
no_c_clust <- c(res_emp$opt_n_clust, res_pharma$opt_n_clust, res_childcare$opt_n_clust, res_health$opt_n_clust, res_grocery$opt_n_clust, res_educpri$opt_n_clust, res_educsec$opt_n_clust, res_lib$opt_n_clust, res_parks$opt_n_clust, res_transit$opt_n_clust)
no_c_clust
```

```{r}
# print("Cut_offs") exp(unique(cutoffs)) - 0.0001
cat("cut_offs_mclust <- list( \n",
  "Emp", (exp(res_emp$breaks) - 0.0001), "\n", "Pharma", exp(res_pharma$breaks) - 0.0001, "\n", "Childcare", exp(res_childcare$breaks) - 0.0001, "\n", "Health", exp(res_health$breaks) - 0.0001, "\n", "Grocery", exp(res_grocery$breaks) - 0.0001, "\n", "Edupri", exp(res_educpri$breaks) -  0.0001, "\n", "Edusec", exp(res_educsec$breaks)- 0.0001, "\n", "Lib", exp(res_lib$breaks)- 0.0001, "\n", "Parks", exp(res_parks$breaks)- 0.0001, "\n", "Transit", exp(res_transit$breaks)- 0.0001)
```


```{r}
cut_offs_jenks <- list(
  PMS_prox_idx_emp = c(0, 0, 0.0003, 0.0011, 0.0032, 0.0083, 0.0194, 0.0441, 0.1076, 0.7249),
  PMS_prox_idx_pharma = c(0.0018, 0.9857),
  PMS_prox_idx_childcare = c(0.001, 0.0079, 0.0155, 0.0267, 0.0429, 0.0668, 0.1035, 0.1654, 0.2941, 0.9208),
  PMS_prox_idx_health = c(0, 0.6759),
  PMS_prox_idx_grocery = c(0.002, 0.7763),
  PMS_prox_idx_educpri = c(0.0095, 0.7452),
  PMS_prox_idx_educsec = c(0.0153, 0.8384),
  PMS_prox_idx_lib = c(0.0166, 0.8665),
  PMS_prox_idx_parks = c(0.0002, 0.823),
  PMS_prox_idx_transit = c(0, 0.5926)
)
```


<!--

```{r eval=False}
Data_emp <- as.matrix(na.omit(master_log$PMS_prox_idx_emp))
emp_breaks <- getJenksBreaks(Data_emp, 4, subset = NULL)
```

```{r eval=False}
emp_breaks
```

```{r eval=False}
emp_breaks_org <- exp(emp_breaks) - 0.0001
emp_breaks_org
```

```{r eval=False}
# add an additional break point
ClusterData_emp <- as.data.frame(Data_emp)
```


```{r eval=False}
# Determine the min and max values
min_val <- min(ClusterData_emp$V1)
max_val <- max(ClusterData_emp$V1)

# Assign group numbers based on breaks
ClusterData_emp$Cluster <- ifelse(ClusterData_emp$V1 <= emp_breaks[1], 1, 
                                  ifelse(ClusterData_emp$V1 < emp_breaks[2], 2,
                                         ifelse(ClusterData_emp$V1 < emp_breaks[3], 3,
                                                ifelse(ClusterData_emp$V1 < emp_breaks[4], 4,
                                                       ifelse(ClusterData_emp$V1 <= max_val, 5, NA)
                                                       )
                                                )
                                         )
                                  )
head(ClusterData_emp)
```


```{r eval=False}
unique(ClusterData_emp$Cluster)
```
-9.2103404 -6.9077553 -4.3428059 -0.3215836

```{r eval=False}
length(ClusterData_emp[ClusterData_emp$V1==-6.9077553,])
```

 
```{r eval=False}
max(ClusterData_emp[ClusterData_emp$Cluster==1, "V1"])
```

```{r eval=False}
min(ClusterData_emp[ClusterData_emp$Cluster==1, "V1"])
```



```{r eval=False}
ClusterData_emp$Cluster <- NA

for (i in 1:length(emp_breaks)) {
  if (i == 1) {
    ClusterData_emp$Cluster <- ifelse(ClusterData_emp$V1 <= emp_breaks[i], i, NA)
  } else {
    ClusterData_emp$Cluster <- ifelse(ClusterData_emp$V1 < emp_breaks[i] & is.na(ClusterData_emp$Cluster), i, ClusterData_emp$Cluster)
  }
}

# Assign the last group to values above the highest break
ClusterData_emp$Cluster <- ifelse(is.na(ClusterData_emp$Cluster), length(emp_breaks) + 1, ClusterData_emp$Cluster)

```


```{r eval=False}
cutoffs <- tapply(ClusterData_emp[, "V1"], ClusterData_emp$Cluster, quantile, probs = c(0, 1))
  
# cut-off viz
# data frame with the cutoff values and group information
cutoff_data <- data.frame(Group = rownames(cutoffs))
cutoff_data$Start <- sapply(cutoffs, function(x) x["0%"])
cutoff_data$End <- sapply(cutoffs, function(x) x["100%"])
```

```{r eval=False}
cutoff_data
```


```{r eval=False}
library(ggplot2)
Data_emp <- data.frame(Value = na.omit(master_sample_log$PMS_prox_idx_emp))
# Create a data frame with the cluster information
cluster_data <- data.frame(Group = c(1, 2, 3, 4, 5),
                           Start = c(-9.2103404, -8.5171932, -6.9077553, -4.3428059, -0.3215836),
                           End = c(-9.2103404, -7.0131158, -4.3505280, -0.3339355, -0.3215836))
```


```{r eval=False}
num_clusters <- length(unique(cluster_data$Group))
cluster_colors <- hcl(h = seq(15, 375, length = num_clusters), c = 100, l = 65)

ggplot(Data_emp) +
  geom_histogram(data = Data_emp, aes(x = Value), alpha = 0.5, bins = 50) + # binwidth = 0.2) +
  geom_density(data = Data_emp,aes(x = Value), alpha = 0.5) +
  # rug cluster
  geom_segment(data = cluster_data, aes(x = Start, xend = End, y = 0, yend = 0, color = as.factor(Group)), size = 2) +
  # rug points
  geom_rug(aes(x = Value), sides = "b", size = 0.2) +
  scale_color_manual(values = cluster_colors) +
  geom_vline(xintercept = cluster_data$Start, color = "blue", linetype = "dashed", size = 0.1, alpha=0.5) +
  geom_vline(xintercept = cluster_data$End, color = "red", linetype = "dashed", size = 0.1, alpha=0.5) +
  ylab("") +
  xlab("Value") +
  ggtitle("Cluster Rugs with Histogram & Cutoff Line") +
  theme_minimal() +
  coord_cartesian(xlim = c(min(cluster_data$Start) - 1, max(cluster_data$End) + 1))

```

```{r eval=False}
num_clusters <- length(unique(cluster_data$Group))
cluster_colors <- hcl(h = seq(15, 375, length = num_clusters), c = 100, l = 65)

ggplot(Data_emp, aes(x = Value)) +
  geom_histogram(aes( y = ..density..),
                 colour = 1, fill = "white", bins = 50) +
  geom_density() +
  # rug cluster
  geom_segment(data = cluster_data, aes(x = Start, xend = End, y = 0, yend = 0, color = as.factor(Group)), size = 2) +
  # rug points
  geom_rug(data = Data_emp, aes(x = Value), sides = "b", size = 0.2) +
  scale_color_manual(values = cluster_colors) +
  geom_vline(data = cluster_data, aes(xintercept = Start), color = "blue", linetype = "dashed", size = 0.1, alpha = 0.5) +
  geom_vline(data = cluster_data, aes(xintercept = End), color = "red", linetype = "dashed", size = 0.1, alpha = 0.5) +
  ylab("") +
  xlab("Value") +
  ggtitle("Cluster Rugs with Histogram & Cutoff Line") +
  theme_minimal() +
  coord_cartesian(xlim = c(min(cluster_data$Start) - 1, max(cluster_data$End) + 1))

```



```{r eval=False}
ClusterData_emp <- as.data.frame(Data_emp)
# Create histogram
hist(ClusterData_emp$V1, breaks = emp_breaks, col = "skyblue", xlab = "Value", ylab = "Frequency",
     main = "Distribution of Data with Jenks Breaks")
```


-->

