---
title: "Manual cutoffs"
author: "Ricky Heinrich"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    extra_dependencies: ["float"]
header-includes:
  \usepackage{placeins}
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = F,  cache = T, message = F, warning = F, fig.height=4, fig.width=6)
# remove float formatting
knitr::opts_chunk$set(echo = F,  cache = T, message = F, warning = F, fig.pos = "H", out.extra = "", fig.align = 'center', fig.height=3, fig.width=6)
library(dplyr)
library(factoextra)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(kableExtra)
```

```{r}
load('../../../data/master_pms_df.Rdata')

# Convert selected columns to numeric
master <- master %>%
  mutate(PMS_DBPOP = as.numeric(gsub(",", "", PMS_DBPOP)), # Dissemination block population
         PMS_DAPOP = as.numeric(gsub(",", "", PMS_DAPOP)), # Dissemination area population
         PMS_CSDPOP = as.numeric(gsub(",", "", PMS_CSDPOP)), # Census subdivision population
         PMS_CMAPOP = as.numeric(gsub(",", "", PMS_CMAPOP)), # Census metropolitan area population
         PMS_PRPOP = as.numeric(gsub(",", "", PMS_PRPOP)), # Province or territory population
         PMS_prox_idx_emp = as.numeric(PMS_prox_idx_emp),
         PMS_prox_idx_pharma = as.numeric(PMS_prox_idx_pharma),
         PMS_prox_idx_childcare = as.numeric(PMS_prox_idx_childcare),
         PMS_prox_idx_health = as.numeric(PMS_prox_idx_health),
         PMS_prox_idx_grocery = as.numeric(PMS_prox_idx_grocery),
         PMS_prox_idx_educpri = as.numeric(PMS_prox_idx_educpri),
         PMS_prox_idx_educsec = as.numeric(PMS_prox_idx_educsec),
         PMS_prox_idx_lib = as.numeric(PMS_prox_idx_lib),
         PMS_prox_idx_parks = as.numeric(PMS_prox_idx_parks),
         PMS_prox_idx_transit = as.numeric(PMS_prox_idx_transit),
         DBUID = as.character(DBUID),
         PMS_DAUID = as.character(PMS_DAUID),
         PMS_CSDUID = as.character(PMS_CSDUID),
         PMS_CMAUID = as.character(PMS_CMAUID),
         PMS_CMAPUID = as.character(PMS_CMAPUID),
         PMS_PRUID = as.character(PMS_PRUID),
         PMS_suppressed = as.character(PMS_suppressed),
         PMS_transit_na = as.character(PMS_suppressed))


# Subset columns that start with "prox_idx"
amenities <- colnames(master)[grepl("^PMS_prox_idx", colnames(master))]
# master dataset - contains only proximity columns
master_amenities <- master[, amenities]

# subsampling data 
perc = 3 #percentage of data to subsample
subsample = (nrow(master)/100)*perc 
master_sample = master[sample(nrow(master), subsample),] #random sample
```
\newpage
# Introduction

The Proximity Measures Database contains continuous measures for 10 amenities for a number of DB within a specific threshold. The distribution of these proximity measures is heavily right skewed, and there are for the most part no discernible clusters. The density distribution, with a default bandwidth, of each amenity is shown in Figure 1.

```{r fig.height=7, fig.width=7, fig.cap="Distribution of proximity measures by amenity"}
master[,amenities] %>% pivot_longer(all_of(amenities)) %>%
  ggplot(aes(value))+ 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 3) + labs(title = "")
```


When transforming the data, the inherent relationship between data points remain the same, but the new structure may reveal new insights. The most 'famous' transformation available is the log transform. It ["can be used to make highly skewed distributions less skewed"](https://stats.libretexts.org/Bookshelves/Introductory_Statistics/Book%3A_Introductory_Statistics_(Lane)/16%3A_Transformations/16.02%3A_Log_Transformations). It may help "make patterns more visible". A consideration to be aware of is that the log of 0 is -Inf. To account for proximity values of 0 in our dataset, we shift the distribution by +0.0001. This avoids the problem of -Inf whilst maintaining the original distances between all values. The downsides of using a log transformation are [DOWNSIDES]. Figure 2 demonstrates the distribution of the log transformed proximity measures, where all the amenities' distributions were shifted by +0.0001. We can already visually identify more possible clusters. 

We shifted the distribution by +0.0001 of the amenities that had a minimum value of 0. Grocery, educpri, educsec, and lib did not have values of 0 in their distribution and such were not shifted. The visual difference of the distributions between when +0.0001 is applied vs when it is not are imperceptible. For simplification in reproducibility, we will apply the distribution shift to all amenities. 

```{r eval = F}
# number of 0s per amenity: expecting 1 per amenity
apply(master_amenities , 2 , function(x) sum(na.omit(x == 0)) )

# minimum proximity measure for each amenity: expecting 0
apply(master_amenities , 2 , function(x) min(na.omit(x)) )

# number of 1s per amenity: expecting 1 per amenity
apply(master_amenities , 2 , function(x) sum(na.omit(x == 1)) )
```


```{r fig.height=7, fig.width=7, fig.cap="LOG TRANSFORMED(0.0001): Distribution of proximity measures by amenity"}
master_log <- master[,amenities] %>% mutate(across(all_of(amenities), ~ .x + 0.0001)) %>% log()

master[,amenities] %>% mutate(across(all_of(amenities), ~ .x + 0.0001)) %>% log() %>%  pivot_longer(all_of(amenities)) %>%
  ggplot(aes(value))+ 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 3) + labs(title = "")
```

```{r fig.height=7, fig.width=7, fig.cap="LOG TRANSFORMED(0.0001 in some): Distribution of proximity measures by amenity", eval = F}
# amenities w/o 0
# amenities[! amenities %in% c('PMS_prox_idx_grocery','PMS_prox_idx_educpri', 'PMS_prox_idx_educsec','PMS_prox_idx_lib' )]

master_log <- master[,amenities] %>% mutate(across(all_of(amenities), ~ .x + 0.0001)) %>% log()

master[,amenities] %>% mutate(across(all_of(amenities[! amenities %in% c('PMS_prox_idx_grocery','PMS_prox_idx_educpri', 'PMS_prox_idx_educsec','PMS_prox_idx_lib' )]), ~ .x + 0.0001)) %>% log() %>%  pivot_longer(all_of(amenities)) %>%
  ggplot(aes(value))+ 
    geom_density() + 
    facet_wrap(~name, scales = 'free_y', ncol = 3) + labs(title = "")
```


\FloatBarrier
# Segmenting via minima

A segmentation technique is to segment the distribution at select minima of the density distribution. Each minimum in the density curves represents a density sparse region, which may be a 'natural' break in the continuous measures. Figure 4 provides an overview of where maxima and minima are located in the density curves of every amenity. We see that there are a lot of points that are by definition local minima, but are not fully indicative of density sparse regions. We can limit which minima are representative of density sparse regions by only including those who have a threshold difference between themselves and surrounding maxima. The results of the indepth analysis of which minima should intuitively represent a cutoff value for each amenity are displayed in the following plot, and summarized in the table.


```{r fig.height=7, fig.width=7, fig.cap="Location of minima and maxima", include = F}
# function to get legend as standalone plot
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  legend
}

## making big list with densities, max/min, plots for each amenity ##
dense_plot_list <- list() # initialize plot list
big_list <- list() # initilize list of amenity lists
count <- 0 

for (amen in amenities){
  # make a list for each amenity, add list to list at end
  # in the list for given amenity, will have density list, max vector, min vector
  count <- count+1
  amen_list <- list()
  amen_list[[1]] <- density(master_log[,amen], na.rm = T) # list of densities
  d_amen <-  density(master_log[,amen], na.rm = T)
  
  # get locations of local minima and maxima
  DeltaY <- diff(d_amen$y) 
  Turns <- which(DeltaY[-1] * DeltaY[-length(DeltaY)] < 0) + 1
  
  # check for cases where have slopes 0
  test_list <- which(DeltaY[-1] * DeltaY[-length(DeltaY)] == 0) 
  test_n <- 0 
  zero_list <- c()
  while (test_n < length(test_list)-1){
    test_n = test_n+1
    if (test_list[test_n+1] != test_list[test_n]+1){
      zero_list <- append(zero_list, test_list[test_n])
    }
  }
  # need to append last one also
  zero_list <- append(zero_list, test_list[length(test_list)])
  Turns <- append(Turns, zero_list) # add to list of turns
  
  
  lst_max <- c()
  lst_min <- c()
  for (i in Turns){
    if (DeltaY[i]<0) {
      lst_max <- append(lst_max, i)
    }
    if (DeltaY[i]>=0){ # assume if slope was 0 then minima
      lst_min <- append(lst_min, i)
    }
  }
  
  lst_min <- sort(lst_min, )
  # deal with rare case where slope is 0 but not actually minima
  # assumes that it's not a maxima or minima, just removes it
  num <- 0
  while (length(lst_min)>= length(lst_max)){ # min list should in theory have 1 less than lst_max
    num <- num + 2
    if (lst_max[num]>lst_min[num]){
      lst_min <- lst_min[-num]
    }
  }
  
  amen_list[[2]] <- lst_max
  amen_list[[3]] <- lst_min
  amen_list[[4]] <- amen_list[[1]][["y"]][amen_list[[2]]]-amen_list[[1]][["y"]][amen_list[[3]]] # diff b/w max and following min density values
  
  # preparing df for plot
  df_dense <- data.frame(x = d_amen$x, y= d_amen$y)
  df_max <- data.frame(x = d_amen$x[lst_max], y = d_amen$y[lst_max], type = "max", index = 1:length(lst_max))
  df_min <- data.frame(x = d_amen$x[lst_min], y = d_amen$y[lst_min], type = "min", index = 1:length(lst_min))
  df_minmax <- rbind(df_max, df_min)
  dense_plot_list[[amen]] <- ggplot(df_dense, aes(x,y))+ geom_line()+geom_point(data = df_minmax, mapping = aes(x,y, col = type))+ theme(legend.position="none") +
    ggtitle(str_sub(amen, 14))
  # add plot to big list
  amen_list[[5]] <- ggplot(df_dense, aes(x,y))+ geom_line()+geom_point(data = df_minmax, mapping = aes(x,y, col = type))+ theme(legend.position="none") +
    ggtitle(str_sub(amen, 14)) # with points
  amen_list[[6]] <- ggplot(df_dense, aes(x,y))+ geom_line()+ theme(legend.position="none") +
    ggtitle(str_sub(amen, 14)) # without points
  amen_list[[7]] <- list(df_minmax)
  
  
  names(amen_list) <- c("density", "maxima", "minima", "maxmindiff", "plot", "nopt", "df_minmax")
  big_list[[count]] <- amen_list # add amenity list to big list
}
names(big_list) <- amenities # rename big list

# set legend: for whichever amen last out of loop, get legend (same for all)
legend_plot <- ggplot(df_dense, aes(x,y))+ geom_line()+geom_point(data = df_minmax, mapping = aes(x,y, col = type))
dense_plot_list[[11]] = g_legend(legend_plot)

# layout of plots
layout_mat <- rbind(c(1:3),
                    c(4:6),
                    c(7:9),
                    c(10:11, 11))
do.call(grid.arrange,list(grobs=dense_plot_list, layout_matrix=layout_mat))

# save image
#cutoffs = do.call(grid.arrange,list(grobs=p, layout_matrix=layout_mat))
#ggsave("cutoffs_hdbscan.png", cutoffs, dpi = 400, width=8, height=5)
```

```{r fig.cap="Employment density curve with minima and maxima", include = F}
# plot
big_list[["PMS_prox_idx_emp"]][["plot"]]
#big_list[["PMS_prox_idx_emp"]][["nopt"]]


# add include row so can do table later
threshold <- 0.001
big_list[[1]][["df_minmax"]][[1]] <- big_list[[1]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[1]][["maxmindiff"]] < threshold), 0,1))
```
```{r fig.cap="Density plot with suggested cutoff points in blue", include=F}
# removing minima that aren't above threshold
# add include row in df
threshold <- 0.001
big_list[[2]][["df_minmax"]][[1]] <- big_list[[2]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[2]][["maxmindiff"]] < threshold), 0,1))



# plot w/ filter
big_list[[2]][["nopt"]] +geom_point(data = filter(big_list[[2]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include=F}
# removing minima that aren't above threshold
# add include row in df
amen_num<- 3
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df
threshold <- 0.001
amen_num = 4 
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df
amen_num = 5
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df
amen_num <- 6
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df
amen_num = 7 
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df

amen_num = 8
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include = F}
# removing minima that aren't above threshold
# add include row in df
amen_num = 9
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

```{r fig.cap="density plot with suggested cutoff points in blue", include=F}
# removing minima that aren't above threshold
# add include row in df
amen_num<- 10
threshold <- 0.001
big_list[[amen_num]][["df_minmax"]][[1]] <- big_list[[amen_num]][["df_minmax"]][[1]] %>% mutate(include = ifelse(index %in% which(big_list[[amen_num]][["maxmindiff"]] < threshold), 0,1))

# plot w/ filter
big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none")
```

## Summary
```{r fig.height=7, fig.width=7}
dense_plot_list_post <- list() 
for (amen_num in 1:length(amenities)){
  dense_plot_list_post[[amen_num]] <- big_list[[amen_num]][["nopt"]] +geom_point(data = filter(big_list[[amen_num]][["df_minmax"]][[1]], include == 1), mapping = aes(x,y, col = type))+ theme(legend.position="none") +
    ggtitle(str_sub(amenities[amen_num], 14))
}

dense_plot_list_post[[11]] = g_legend(legend_plot)

# layout of plots
layout_mat <- rbind(c(1:3),
                    c(4:6),
                    c(7:9),
                    c(10:11, 11))
do.call(grid.arrange,list(grobs=dense_plot_list_post, layout_matrix=layout_mat))

```

```{r}
# remade this to have reproducible workflow instead of manual count
# the number of maxima gives essentially the number of groups
# sum the 'include' col in df_minmax if type is max for each amenity
# ezpz
num_groups <- c()
for (ameni in 1:length(amenities)){
  num_groups[ameni] <- aggregate(include~type, data = big_list[[ameni]][["df_minmax"]][[1]], sum)[1,2]
}


kable(cbind(str_sub(amenities, 14), num_groups)) %>% kable_styling(latex_options = "HOLD_position")
```

Cut off values:
```{r }
# cluster values
# for each amenity get x-values of minima
# so each amenity has a list
# in big_list df_minmax, if type is min AND include is 1, then get the density x-value
# all in same df_minmax ezpz
for (ameni in 1:length(amenities)){
  print(amenities[ameni])
  print(exp(filter(big_list[[ameni]][["df_minmax"]][[1]], type == "min" & include == 1)$x))
}
```
```{r eval = F}
# i guess have to manually make this list ... ran the above for loop and copied the values here
num_of_clust_manual <- c(5, 3, 3, 4, 3, 3, 2, 2, 3, 5)

cut_offs_manual <- list(
  PMS_prox_idx_emp = c(0.0001423603, 0.0002573125, 0.0003743458, 0.0006010775),
  PMS_prox_idx_pharma = c(0.01152592, 0.01958168),
  PMS_prox_idx_childcare = c(0.008507409, 0.013950871),
  PMS_prox_idx_health = c(0.0001405736, 0.0002524638, 0.0003658077),
  PMS_prox_idx_grocery = c(0.01218361, 0.01856220),
  PMS_prox_idx_educpri = c(0.04975835, 0.08179011),
  PMS_prox_idx_educsec = c(0.06619424),
  PMS_prox_idx_lib = c(0.6150259),
  PMS_prox_idx_parks = c(0.01844893, 0.02954933),
  PMS_prox_idx_transit = c(0.0001390986, 0.0002481958, 0.0003512994, 0.0004514904, 0.8553858919)
)
```

Essentially, for every amenity, add a new column (so 10 new cols today) that outlines which group it belongs to, according to the cutoff points.

```{r}
# set the group for each prox value in each amenity: add 10 new cols
for (ameni in 1:length(amenities)){
  cutoffs <- c(filter(big_list[[ameni]][["df_minmax"]][[1]], type == "min" & include == 1)$x)
  min_x <- big_list[[ameni]][["density"]][["x"]][1]
  max_x <- big_list[[ameni]][["density"]][["x"]][length(big_list[[ameni]][["density"]][["x"]])]
  
  # vector: min, cutoffs, max
  group_boundaries <- c(min_x, cutoffs, max_x)
  
  # new col with interval val belongs to
  master_log[,paste0("group",ameni)]  <- findInterval(master_log[,ameni], group_boundaries)
}
```

```{r fig.cap="Table of DBs per group for Emp"}
# summary stats

# gives # of DBs per group
#table(master_log$group1)
# make df of table
# combine df of other tables
kable(t(table(master_log$group1,useNA = "ifany"))) %>% kable_styling(latex_options = "HOLD_position")

df_DBcount <- data.frame(table(master_log$group1,useNA = "ifany"))
df_DBcount2 <- data.frame(table(master_log$group2,useNA = "ifany"))
df_transDB <- t(df_DBcount)
df_transDB2 <- t(df_DBcount2)

header.true <- function(df) {
  names(df) <- as.character(unlist(df[1,]))
  df[-1,]
}

df_transDB <- header.true(df_transDB)
```

## How many DBs in each group

```{r}
# cols that start with 'group
grupscols <- colnames(master_log)[grepl("^group", colnames(master_log))]

n=0
for (grup in grupscols){
  n = n+1
  print(str_sub(amenities[n], 14))
  print(t(table(master_log[,grup],useNA = "ifany")))
}

```

## Non-logged: adding to master df
```{r}
for (ameni in 1:length(amenities)){
  cutoffs <- exp(c(filter(big_list[[ameni]][["df_minmax"]][[1]], type == "min" & include == 1)$x))-0.0001
  
  min_x <- 0
  max_x <-  1.1
  
  # vector: min, cutoffs, max
  group_boundaries <- c(min_x, cutoffs, max_x)
  
  # new col with interval val belongs to
  master[,paste0("group",ameni)]  <- findInterval(master[,amenities[ameni]], group_boundaries)
}

# cols that start with 'group
grupscols <- colnames(master)[grepl("^group", colnames(master))]

# table(master$group1,useNA = "ifany")

n=0
for (grup in grupscols){
  n = n+1
  print(str_sub(amenities[n], 14))
  print(t(table(master[,grup],useNA = "ifany")))
}
```

```{r}
exp(group_boundaries)-0.0001
max(na.omit(master$PMS_prox_idx_emp))
```


## summary statistics of IoR



# Next Steps
- ~~assign 'group' to each~~
- metrics for each clusters
- Investigate 'weight' of distribution for every suggested groups
- ~~number of DB for each clusters~~
- summary statistics of IoR for each cluster
- population summary statistics
- CMA type summary statistics

